---
title: 'STAT 687 Homework #5'
output: 
  pdf_document:
    fig_height: 4
    fig_width:  5
---

```{r echo=FALSE,message=FALSE}
library(meekR)
library(survival)
library(actuar)
library(evd)
```

# From the text problem # 15.2

_Revisit the shock absorber data introduced in Example 3.8 and given in Appendix Table C.2.  There were actually two different causes of failure for this product.  The analyses in Chapters 3 and 8 ignored the different "failure modes" in the analysis.  An alternative analysis (important for some purposes) would take into consideration the different failure modes.  Such an analysis is straightforward when the failure times for the two different modes are statistically independent._

(a) _Assuming that the failure modes are independent, fit Weibull distributions separately to estimate the time to failure for each mode.  Also fit other parametric distributions to the individual failure modes to see if there are other distributions fit better than the Weibull._

(b) _Suppose that the shock absorber is a series system that fails as soon as there is a failure due to one mode or the other.  Combine the results from part (a) to obtain an estimate of the cdf for the shock absorbers with both modes acting together.  Plot these on Weibull probability paper and compare the results with the analyses that ignores the differences between the two different failure modes._

(c) _Provide an intuitive explanation for the result in part (b)_

(d) _Explain why the agreement between the two methods of analysis is so good in this case (as compared to Example 15.6). In general, when would you expect to see more important differences between the analysis that accounts for the different failure modes and the analysis that does not account for the different failure modes?_

# From the text problem # 17.3

_McCool (1980) gives the result of a life test on rolling contact fatigue of ceramic ball bearings.  Ten specimens were tested at each of four levels of stress.  The ordered failure times are given in the following table.  McCool indicates that it is customary to model such data with the two-parameter Weibull distribution with a shape parameter that does not depend on stress._

(a) _Plot the failure times versus stress on log-log axes (or, alternately take logs and plot on linear axes)._

(b) _It is often suggested that median failure time is proportional to a power transformation of stress: that is,_ $t_{.5}=e^{\beta_{0}}\times(\text{stress})^{\beta_{1}}$ _or_ $\log(t_{.5})=\beta_{0}+\beta_{1}\text{log(stress)}.$  _Is the suggestion reasonable in this case?  Plot the sample medians on the graph in part (a) to help answer this question._

(c) _Use a hand drawn line through the sample versus stress points to obtain a graphical estimate of the exponent (or slope)_ $\beta_{1}$.

(d) _Make separate Weibull probability plots for the data at each level of stress, plotting them all on the same graph.  What does this plot suggest about the customary assumption that the Weibull shape parameter is the same for all levels of stress?  Provide possible explanations for the observed differences in the estimates of the Weibull shape parameter at each level of stress_

# From the text problem # 17.5

_Refer to Exercise 17.3.  Suppose that log life can be described adequately by a normal distribution.  Consider the regression model_ $\mu=\beta_0+\beta_1\times\text{log}(stress)$ _with constant_ $\sigma,$ _where_ $\sigma$ _is the standard deviation of log life, the same for any fixed level of stress._

(a) _Use ordinary least squares to fit this model to the rolling contact fatigue data.  Compare your answer to the graphical estimate from part (c) of Exercise 17.3._

(b) _Use a computer that does maximum likelihood estimation and fit the lognormal regression model (17.7) to the rolling contact fatigue data_

(c) _Compute and compare estimates from these two different methods of estimation for the median time to failure at_ $1.05\times10^{6}$ psi, _the_ $.01$ _quantile at_ $1.05 \times 10^{6}$ _psi and the_ $.01$ _quantile at_ $0.85\times10^{6}$ _psi.  Comment on the results of this comparison._

# From the text problem # 18.2

_Based on previous experience with similar products, the failure time of a particular field effect transistor on a monolithic microwave GaAs integrated circuit, operating at $100^{o}C$ (channel temperature), is expected to have a lognormal distribution with a median time to failure of $30$ years.  The  primary failure mode is caused by a chemical reaction that has an activation energy of_ $E_{a}=.6 eV.$ _The value of the lognormal scale parameter for this failure mode is expected to be $\sigma=.7$._

(a) _What are the lognormal parameters $\mu$ and $\sigma$ if time is recorded in hours?_

(b) _For operation at_ $100^{o}C$ _channel temperature, what is the time at which $5\%$ of the units would fail?_ $10\%?, 90\%$?

(c) _Calculate the time-acceleration factor for testing at_ $250^{o}C$, $200^{o}C,$ _and_ $150^{o}C$ _channel temperature.  Use Table 18.2 and Figure 18.3 and check with equation (18.2)._

(d) _Obtain an expression for the temperature at which $100p\%$ of tested units would be expected to fail in a_ $6000$ _hour test.  Use this expression to compute the temperatures at which_ $90\%$ _and_ $10\%$ _would be expected to fail._

# From the text problem # 18.3

_Refer to Exercise 18.2.  Obtain an expression for the (average) FIT rate (in standard units of failures per hour in parts per billion) for the first 10 years of operation at_ $100^{o}C$.  _How much would this improve if the operating channel temperature is changed to only_ $90^{o}C?$

# From the text problem # 19.2

_The following table contains data from an accelerated life test on Device-C, an integrated circuit.  Failures were caused by a chemical reaction inside the circuit package.  Reliability engineers tested 10 circuits at each temperature over a period of_ $3000$ _hours.  The purpose of the experiment was to estimate the activation energy of the failure causing reaction and to obtain an estimate of the integrated circuit life distribution at_ $80^{o}C$ _junction temperature._

```{r echo=FALSE,results='asis',message=FALSE}
library(xtable)
Jtemp<-c("80 C","125 C","150 C","175 C","200 C")
ntest<-c(rep(10,5))
Fail <-c("None by 3000 hours","None by 3000 hours","2.35,2.56,2.98",".80,1.13,1.21,1.31,1.35,1.37,1.42,1.77,1.96",".22,.25,.28,.33,.38,.46,.46,.51,.61")
dat.f<-data.frame(Jtemp,ntest,Fail)
names(dat.f)<-c("Junction Temperature", "# of Units Tested","Recorded Failure Times in Thousands of Hours")
dat.fx<-xtable(dat.f,align=c("c","c","c","l"))
print(dat.fx,floating=TRUE,comment=FALSE)
```

(a) _For each temperature with failures, plot the ordered failure times_ $t_{(i)}$ _versus_ $(i-.5)/10$ _on lognormal probability paper_
(b) _Repeat part (a) but use Weibull probability paper_
(c) _Make a judgement as to whether the lognormal or the Weibull distribution is a more adequate distribution for these data_
(d) _Using the lognormal probabilty plot, draw a set of parallel straight lines through the plotted points, one line for each temperature having failures_
(e) _Use each line on the probability plot to obtain graphical estimates of the $.5$ quantiles at the corresponding temperatures._
(f) _Use each line on the probability plot to obtain graphical estimates of the $.01$ quantiles at the corresponding temperatures.  Describe the nature of the extrapolation in these estimates_
(g) _Plot the estimates of the_ $.5$ _quantile versus temperature on Arrhenius paper.  Draw a straight line to estimate the relationship between life and temperature.  Do the same with the graphical estimates of the_ $.01$ _quantile._
(h) _Graphically estimate the slope of the lines drawn in part (g), and use these to obtain a graphical estimate of the failure mode's activation energy._
(i) _Use the lines drawn in part (g) to obtain estimates of the_ $.01$ _and the_ $.5$ _quantiles of the life distribution of Device-C at_ $80^{o}C$.  _Describe the nature of the extrapolation in these estimates._
(j) _Predict the effect on the estimates of the Arrhenius relationship in the data at_ $80^{o}C$ _and_ $125^{o}C$ _were omitted from the analysis.  Explain._