---
title: 'STAT 687 - Homework # 5'
author: "Due 0800, 21 July 2014"
fontsize: 11in
output:
  pdf_document:
    fig_cap: yes
    includes:
      in_header: ../css/jkf.header.tex
  html_document:
    css: ../css/homework.css
    fig_cap: yes
    highlight: pygments
    theme: readable
    toc: yes
geometry: margin=1in
citation_package: natbib
---

```{r,echo=FALSE,message=FALSE, warning=FALSE}
source('../css/setup.R')
output <- getYAML('stat687-hw5.Rmd')$output
if(!    'SMRD'%in%library()$results) install.packages('SMRD')     ; library(SMRD)
if(!'survival'%in%library()$results) install.packages('survival') ; library(survival)
if(!  'actuar'%in%library()$results) install.packages('actuar')   ; library(actuar)
if(!     'evd'%in%library()$results) install.packages('evd')      ; library(evd)
```

## From the text problem # 15.2

Revisit the shock absorber data introduced in Example 3.8 and given in Appendix Table C.2.  There were actually two different causes of failure for this product.  The analyses in Chapters 3 and 8 ignored the different "failure modes" in the analysis.  An alternative analysis (important for some purposes) would take into consideration the different failure modes.  Such an analysis is straightforward when the failure times for the two different modes are statistically independent.

(a) Assuming that the failure modes are independent, fit Weibull distributions separately to estimate the time to failure for each mode.  Also fit other parametric distributions to the individual failure modes to see if there are other distributions fit better than the Weibull.

(b) Suppose that the shock absorber is a series system that fails as soon as there is a failure due to one mode or the other.  Combine the results from part (a) to obtain an estimate of the cdf for the shock absorbers with both modes acting together.  Plot these on Weibull probability paper and compare the results with the analyses that ignores the differences between the two different failure modes.

(c) Provide an intuitive explanation for the result in part (b)

(d) Explain why the agreement between the two methods of analysis is so good in this case (as compared to Example 15.6). In general, when would you expect to see more important differences between the analysis that accounts for the different failure modes and the analysis that does not account for the different failure modes?

`r if(output=='pdf') '\\newpage'`

## From the text problem # 17.3

McCool (1980) gives the result of a life test on rolling contact fatigue of ceramic ball bearings.  Ten specimens were tested at each of four levels of stress.  The ordered failure times are given in the following table.  McCool indicates that it is customary to model such data with the two-parameter Weibull distribution with a shape parameter that does not depend on stress.

(a) Plot the failure times versus stress on log-log axes (or, alternately take logs and plot on linear axes).

(b) It is often suggested that median failure time is proportional to a power transformation of stress: that is, $t_{.5}=e^{\beta_{0}}\times(\text{stress})^{\beta_{1}}$ or $\log(t_{.5})=\beta_{0}+\beta_{1}\text{log(stress)}.$  Is the suggestion reasonable in this case?  Plot the sample medians on the graph in part (a) to help answer this question.

(c) Use a hand drawn line through the sample versus stress points to obtain a graphical estimate of the exponent (or slope) $\beta_{1}$.

(d) Make separate Weibull probability plots for the data at each level of stress, plotting them all on the same graph.  What does this plot suggest about the customary assumption that the Weibull shape parameter is the same for all levels of stress?  Provide possible explanations for the observed differences in the estimates of the Weibull shape parameter at each level of stress

`r if(output=='pdf') '\\newpage'`

## From the text problem # 17.5

Refer to Exercise 17.3.  Suppose that log life can be described adequately by a normal distribution.  Consider the regression model $\mu=\beta_0+\beta_1\times\text{log}(stress)$ with constant $\sigma,$ where $\sigma$ is the standard deviation of log life, the same for any fixed level of stress.

(a) Use ordinary least squares to fit this model to the rolling contact fatigue data.  Compare your answer to the graphical estimate from part (c) of Exercise 17.3.

(b) Use a computer that does maximum likelihood estimation and fit the lognormal regression model (17.7) to the rolling contact fatigue data

(c) Compute and compare estimates from these two different methods of estimation for the median time to failure at $1.05\times10^{6}$ psi, the $.01$ quantile at $1.05 \times 10^{6}$ psi and the $.01$ quantile at $0.85\times10^{6}$ psi.  Comment on the results of this comparison.

`r if(output=='pdf') '\\newpage'`

## From the text problem # 18.2

Based on previous experience with similar products, the failure time of a particular field effect transistor on a monolithic microwave GaAs integrated circuit, operating at $100^{o}C$ (channel temperature), is expected to have a lognormal distribution with a median time to failure of $30$ years.  The  primary failure mode is caused by a chemical reaction that has an activation energy of $E_{a}=.6 eV.$ The value of the lognormal scale parameter for this failure mode is expected to be $\sigma=.7$.

(a) What are the lognormal parameters $\mu$ and $\sigma$ if time is recorded in hours?

(b) For operation at $100^{o}C$ channel temperature, what is the time at which $5\%$ of the units would fail? $10\%?, 90\%$?

(c) Calculate the time-acceleration factor for testing at $250^{o}C$, $200^{o}C,$ and $150^{o}C$ channel temperature.  Use Table 18.2 and Figure 18.3 and check with equation (18.2).

(d) Obtain an expression for the temperature at which $100p\%$ of tested units would be expected to fail in a $6000$ hour test.  Use this expression to compute the temperatures at which $90\%$ and $10\%$ would be expected to fail.

`r if(output=='pdf') '\\newpage'`

## From the text problem # 18.3

Refer to Exercise 18.2.  Obtain an expression for the (average) FIT rate (in standard units of failures per hour in parts per billion) for the first 10 years of operation at $100^{o}C$.  How much would this improve if the operating channel temperature is changed to only $90^{o}C?$

`r if(output=='pdf') '\\newpage'`

## From the text problem # 19.2

The following table contains data from an accelerated life test on Device-C, an integrated circuit.  Failures were caused by a chemical reaction inside the circuit package.  Reliability engineers tested 10 circuits at each temperature over a period of $3000$ hours.  The purpose of the experiment was to estimate the activation energy of the failure causing reaction and to obtain an estimate of the integrated circuit life distribution at $80^{o}C$ junction temperature.

```{r echo=FALSE,results='asis',message=FALSE}
library(xtable)
Jtemp<-c("80 C","125 C","150 C","175 C","200 C")
ntest<-c(rep(10,5))
Fail <-c("None by 3000 hours","None by 3000 hours","2.35,2.56,2.98",".80,1.13,1.21,1.31,1.35,1.37,1.42,1.77,1.96",".22,.25,.28,.33,.38,.46,.46,.51,.61")
dat.f<-data.frame(Jtemp,ntest,Fail)
names(dat.f)<-c("Junction Temperature", "# of Units Tested","Recorded Failure Times in Thousands of Hours")
dat.fx<-xtable(dat.f,align=c("c","c","c","l"))
print(dat.fx,floating=TRUE,comment=FALSE)
```

(a) For each temperature with failures, plot the ordered failure times $t_{(i)}$ versus $(i-.5)/10$ on lognormal probability paper
(b) Repeat part (a) but use Weibull probability paper
(c) Make a judgement as to whether the lognormal or the Weibull distribution is a more adequate distribution for these data
(d) Using the lognormal probabilty plot, draw a set of parallel straight lines through the plotted points, one line for each temperature having failures_
(e) Use each line on the probability plot to obtain graphical estimates of the $.5$ quantiles at the corresponding temperatures.
(f) Use each line on the probability plot to obtain graphical estimates of the $.01$ quantiles at the corresponding temperatures.  Describe the nature of the extrapolation in these estimates
(g) Plot the estimates of the $.5$ quantile versus temperature on Arrhenius paper.  Draw a straight line to estimate the relationship between life and temperature.  Do the same with the graphical estimates of the $.01$ quantile.
(h) Graphically estimate the slope of the lines drawn in part (g), and use these to obtain a graphical estimate of the failure mode's activation energy.
(i) Use the lines drawn in part (g) to obtain estimates of the $.01$ and the $.5$ quantiles of the life distribution of Device-C at $80^{o}C$.  Describe the nature of the extrapolation in these estimates.
(j) Predict the effect on the estimates of the Arrhenius relationship in the data at $80^{o}C$ and $125^{o}C$ were omitted from the analysis.  Explain.