%references for Meeker and Escobar
%original by wqmeeker 27 mar 1994
%edited by wqmeeker 19 nov 1996
\begin{thebibliography}{999}

\bibitem{aalen76} 
Aalen, O. (1976), Nonparametric inference in connection with multiple
decrement models, {\em Scandinavian Journal of Statistics}, {\bf 3}, 15--27.

\bibitem{ah91} 
Aalen, O., and Husebye, E. (1991), Statistical analysis of repeated
events forming renewal processes, {\em Statistics in Medicine}, {\bf
10}, 1227--1240.

\bibitem{96}
Abernethy, R. B. (1996), {\em The New Weibull Handbook} (Second
Edition), published by Robert B. Abernethy, 536 Oyster Road, North
Palm Beach, FL 33408--4328.

\bibitem{abmr83}
Abernethy, R. B., Breneman, J. E., Medlin, C. H., and Reinman, G. L.
(1983), {\em Weibull Analysis Handbook}, Air Force Wright Aeronautical
Laboratories Technical Report AFWAL-TR-83-2079. Available from the
National Technical Information Service, Washington, DC.

\bibitem{akritas86}
Akritas, M. G. (1986), Bootstrapping the Kaplan--Meier estimator, {\em
Journal of the American Statistical Association}, {\bf 81}, 1032--1038.

\bibitem{ab82}
Amster, S. J., Brush, G. G., and Saperstein, B. (1982), Planning
and conducting field-tracking studies, {\em The Bell System
Technical Journal}, {\bf 61}, 2333--2364.

\bibitem{ah83}
Amster, S. J., and Hooper, J. H. (1983), Accelerated life tests with
measured degradation data and growth curve models, paper presented
at the American Statistical Association Annual Meetings, Toronto.

\bibitem{abgk93}
Anderson, P. K., Borgan, \O., Gill, R. D., and Keiding, N. (1993),
{\em Statistical Models Based on Counting Processes}, New York:
Springer-Verlag.

\bibitem{anx64}
Anscombe, F. J. (1964), Normal likelihood functions, {\em Annals of
the Institute of Statistical Mathematics} {\bf 16}, 1--19.

\bibitem{ap89}
Ansell, J. I., and Phillips, M. J. (1989), Practical problems in the
statistical analysis of reliability data, {\em Applied Statistics},
{\bf 38}, 205--231.

\bibitem{ap94}
Ansell J. I., and Phillips, M. J. (1994), {\em Practical Methods for
Reliability Data Analysis}, Oxford: Clarendon Press.

\bibitem{af84}
Ascher, H., and Feingold, H. (1984),
{\em Repairable Systems Reliability}, New York: Marcel Dekker.

\bibitem{atwood84}
Atwood, C. L. (1984), Approximate tolerance intervals, based on
maximum likelihood estimates, {\em Journal of the American Statistical
Association}, {\bf 79}, 459--465.


\bibitem{az95}
Azem, A. (1995), {\em Software Reliability Determination for
Conventional and Logic Programming},, New York: Walter de Gruyter.


\bibitem{bn95}
Bagdonavi\v{c}ius, V., and Nikulin, M. (1995), Semiparametric models in
accelerated life testing, {\em Queen's papers in Pure and Applied
Mathematics}, {\bf 98}, Kingston: Queen's University.

\bibitem{bn97}
Bagdonavi\v{c}ius, V., and Nikulin, M. (1997), Additive and
multiplicative semiparametric models in accelerated life testing and
survival analysis, {\em Queen's Papers in Pure and Applied
Mathematics}, {\bf xxx}, Kingston: Queen's University.

\bibitem{bg81}
Bailey, R. A., and Gilbert, R. A. (1981), STRIFE testing for
reliability improvement, {\em Proceedings of the Institute of
Environmental Sciences}, 119--121.

\bibitem{be91}
Bain, L. J., and Engelhardt, M. (1991), {\em Statistical Analysis of
Reliability and Life Testing Models, Theory and Methods} (Second
Edition), New York: Marcel Dekker.

\bibitem{bala91}
Balakrishnan, N., Editor (1991), {\em The Logistic Distribution}, New
York: Marcel Dekker.

\bibitem{bp75}
Barlow, R. E., and Proschan, F. (1975), {\em Statistical Theory of
Reliability and Life Testing}, New York: Holt, Rinehart, and Winston.

\bibitem{barnet76}
Barnett, V. (1976), Convenient plotting positions for the normal
distribution, {\em Applied Statistics}, {\bf 25}, 47--50.

\bibitem{bt94}
Baxter, L. A., and Tortorella, M. (1994), Dealing with real field
reliability data: circumventing incompleteness by modeling \&
iteration.  {\em 1994 Proceedings Annual Reliability and
Maintainability Symposium}, 225--262, New York: Institute of
Electrical and Electronics Engineers.

\bibitem{bayer94}
Bayer, R. G. (1994), {\em Mechanical Wear Prediction and Prevention},
New York: Marcel Dekker.

\bibitem{beckwith79}
Beckwith, J. P. (1979), Estimation of the strength remaining of a
material that decays with time, Department of Mathematics and Computer
Science, Michigan Technological University, Houghton, MI.

\bibitem{beckwith80}
Beckwith, J. P. (1980), An estimator and design technique for the
estimation of a rate parameter in accelerated testing, Department of
Mathematics and Computer Science, Michigan Technological University,
Houghton, MI.


\bibitem{bendell85}
Bendell, A. (1985), Proportional hazards modeling in reliability
assessment, {\em Reliability Engineering}, {\bf 11}, 1975--1983.

\bibitem{beran90} 
Beran R. (1990), Calibrating prediction regions, {\em Journal of the
American Statistical Association}, {\bf 85}, 715--723.

\bibitem{berkson66}
Berkson, J. (1966), Examination of randomness of $\alpha$-particle
emissions, in {\em Festschrift for J. Neyman, Research Papers in
Statistics}, F. N. David, Editor. New York: John Wiley \& Sons.

\bibitem{bhatt85}
Bhattacharyya, G. K. (1985), The asymptotics of maximum likelihood and
related estimators based on type II censored data, {\em Journal of the
American Statistical Association}, {\bf 80}, 398--404.

\bibitem{billingsley86}
Billingsley, P. (1986), {\em Probability and Measure} (Second
Edition), New York: John Wiley \& Sons.

\bibitem{birn79}
Birnbaum, Z. W. (1979), {\em On the Mathematics of Competing Risks},
Department of Health, Education and Welfare Publication (PHS)79-1351.
Available from the Superintendent of Documents, U.S.  Government
Printing Office, Washington, DC 20402.

\bibitem{bs69}
Birnbaum, Z. W., and Saunders, S. C. (1969), A new family of life
distributions, {\em Journal of Applied Probability}, {\bf 6}, 319--327.

\bibitem{black69}
Black, J. R. (1969), Electromigration---a brief survey and some
recent results, {\em IEEE Transactions on Electronic Devices}, {\bf
ED-16}, 338--347.


\bibitem{bm96}
Blischke, W. R., and Murthy, D. N. P., Editors (1996), {\em Product Warranty
Handbook}, New York: Marcel Dekker.

\bibitem{blom58}
Blom, G. (1958), {\em Statistical Estimates and Transformed
Beta-Variables}, New York: John Wiley \& Sons.

\bibitem{bgh73}
Blumenthal, S., Greenwood, J. A., and Herbach, L. (1973), The transient
reliability behavior of series systems or superimposed renewal
processes, {\em Technometrics}, {\bf 15}, 255--269.

\bibitem{bgh76}
Blumenthal, S., Greenwood, J. A., and Herbach, L. (1976),
A comparison of the bad as old and superimposed renewal models,
{\em Management Science}, {\bf 23}, 280--285.


\bibitem{bbdg89}
Boccaletti, G., Borri, F. R., D'Esponosa, F., and Ghio, E.
(1989), Accelerated Tests, Chapter 11 in {\em Microelectronic
Reliability, Volume II, Reliability, Integrity Assessment and
Assurance}, E. Pollino, Editor.  Norwood, MA: Artech House.

\bibitem{bk85}
Bogdanoff, J. L., and Kozin, F. (1985), {\em Probabilistic Models of
Cumulative Damage}, New York: John Wiley \& Sons.

\bibitem{be94}
Boulanger, M., and Escobar, L. A. (1994), Experimental design for a
class of accelerated degradation tests,  {\em Technometrics}, {\bf 36},
260--272.

\bibitem{bd87}
Box, G. E. P., and Draper, N. R. (1987), {\em Empirical Model-Building and
Response Surfaces}, New York: John Wiley \& Sons.

\bibitem{bt73}
Box, G. E. P., and Tiao, G. C. (1973), {\em Bayesian Inference in
Statistical Analysis}, Reading, MA: Addison-Wesley.

\bibitem{bk89}
Boyko, K. C., and Gerlach, D. L. (1989), Time dependent dielectric
breakdown of 210 \AA~oxides, {\em Proceedings of the IEEE International
Reliability Physics Symposium}, {\bf 27}, 1--8.

\bibitem{bl90}
Bro, P., and Levy, S. C. (1990), {\em Quality and Reliability Methods
for Primary Batteries}, New York: John Wiley \& Sons.

\bibitem{bm79}
Brown, H. M., and Mains D. E (1979), {\em Accelerated Test Program for
Sealed Nickel--Cadmium Spacecraft Batteries/Cells},
Technical Report WQEC/C 79-145. Available from the Department of the
Navy, Naval Weapons Support Center, Weapons Quality Engineering
Center, Crane, IN 47522.

\bibitem{brownlee60}
Brownlee, K. A. (1960), {\em Statistical Theory and Methodology in
Science and Engineering}, New York: John Wiley \& Sons.


\bibitem{buckland85}
Buckland, S. T. (1985), Calculation of Monte Carlo confidence intervals,
{\em Applied Statistics}, {\bf 34}, 296--301.

\bibitem{bq93}
Byrne, D., and Quinlan, J. (1993), Robust function for attaining high
reliability at low cost, {\em 1993 Proceedings Annual Reliability and
Maintainability Symposium}, 183--191, New York: Institute of
Electrical and Electronics Engineers.

\bibitem{carey89}
Carey, M. B. (1989), Challenges of reliability assessment based on
degradation data: an example, paper presented at the American
Statistical Association 150th Annual Meeting, Washington, DC.

\bibitem{ck91}
Carey, M. B., and Koenig, R. H. (1991), Reliability assessment based
on accelerated degradation: a case study,  {\em IEEE Transactions on
Reliability}, {\bf 40}, 499--506.

\bibitem{ct88}
Carey, M. B., and Tortorella, M. (1988), Analysis of degradation
data applied to MOS devices, paper presented at the 6th
International Conference on Reliability and Maintainability,
Strasbourg, France.

\bibitem{cb90}
Casella, G., and Berger, R. L. (1990), {\em Statistical Inference},
Belmont, CA: Wadsworth.

\bibitem{castillo88}
Castillo, E. (1988), {\em Extreme Value Theory in Engineering},
New York: Academic Press.

\bibitem{cg87}
Castillo, E., and Galambos, J. (1987), Lifetime regression models
based on a functional equation of physical nature, {\em Journal of
Applied Probability}, {\bf 24}, 160--169.

\bibitem{cl92}
Chaloner, K., and Larntz, K. (1992), Bayesian design for accelerated
life testing, {\em Journal of Statistical Planning and Inference},
{\bf 33}, 245--259.

\bibitem{cbt94}
Chan, C. K., Boulanger, M., and Tortorella, M. (1994), Analysis of
parameter-degradation data using life-data analysis programs, {\em
1994 Proceedings Annual Reliability and Maintainability Symposium},
228--291, New York: Institute of
Electrical and Electronics Engineers.

\bibitem{cm97}
Chan, V., and Meeker, W. Q. (1998), A competing-risk limited failure
population model for product failure times, Department of
Statistics, Iowa State University, Ames, IA.

\bibitem{chang92}
Chang, D. S. (1992), Analysis of accelerated degradation data in a
two-way design, {\em Reliability Engineering and System Safety},
{\bf 39}, 65--69.

\bibitem{ci83}
Cheng, R. C. H., and Iles, T. C. (1983), Confidence bands for
cumulative distribution functions of continuous random variables,
{\em Technometrics}, {\bf 25}, 77--86.

\bibitem{ci87}
Cheng, R. C. H., and Iles, T. C. (1987), Corrected maximum likelihood
in non-regular problems, {\em Journal of the Royal Statistical
Society B}, {\bf 49}, 95--101.

\bibitem{ci88}
Cheng, R. C. H., and Iles, T. C. (1988), One-sided confidence bands
for cumulative distribution functions, {\em Technometrics}, {\bf
30}, 155--159.

\bibitem{ci90}
Cheng, R. C. H., and Iles, T. C. (1990), Embedded models in
three-parameter distributions and their estimation, {\em Journal of
the Royal Statistical Society B}, {\bf 52}, 135--149.

\bibitem{cl54}
Chernoff, H., and Lieberman, G. J.  (1954), Use of normal
probability paper, {\em Journal of the American Statistical
Association}, {\bf 49}, 778--785.

\bibitem{cf98}
Chhikara, R. S., and Folks, J. L. (1989), {\em The Inverse Gaussian
Distribution}, New York: Marcel Dekker.

\bibitem{cl95}
Chow, S. C., and Liu, J. P. (1995), {\em Statistical Design and
Analysis in Pharmaceutical Science: Validation, Process Controls,
and Stability}, New York: Marcel Dekker.

\bibitem{christou92} 
Christou, A. (1992), {\em Reliability of Gallium Arsenide MMICs},
New York: John Wiley \& Sons.

\bibitem{christou94a} 
Christou, A. (1994a), {\em Integrating Reliability into
Microelectronics Manufacturing}, New York: John Wiley \& Sons.

\bibitem{christou94b} 
Christou, A. (1994b), {\em Electromigration and Related Electronic Device
Degradation}, New York: John Wiley \& Sons.

\bibitem{cohen91}
Cohen, A. C. (1991),
{\em Truncated and Censored Samples. Theory and Applications},
New York:  Marcel Dekker.

\bibitem{cw88}
Cohen, A. C., and Whitten, B. J. (1988),
{\em Parameter Estimation in Reliability and Life Span Models},
New York:  Marcel Dekker.

\bibitem{collett}
Collett, D. (1994),
{\em Modelling Survival Data in Medical Research}, New York:
Chapman \& Hall.

\bibitem{condra93}
Condra, L. W.  (1993), {\em Reliability Improvement with Design of
Experiments}, New York: Marcel Dekker.

\bibitem{cct91}
Confer, R., Canner, J., and Trostle, T. (1991), Use of highly
accelerated life test (HALT) to determine reliability of multilayer
ceramic capacitors, {\em Proceedings of the Electronic Components \&
Technology Conference}, {\bf 41}, 320--322.

\bibitem{cm93}
Costa, J. M., and Mercer, A. D. (1993), {\em Progress in the
Understanding and Prevention of Corrosion}, London: The Institute of
Materials.

\bibitem{cox62}
Cox, D. R.  (1962), {\em Renewal Theory}, London: Methuen.

\bibitem{cox75}
Cox, D. R. (1975), Prediction intervals and empirical Bayes
confidence intervals, in {\em Perspectives in Probability and
Statistics}, J. Gani, Editor, London: Academic Press.

\bibitem{ch74}
Cox, D. R., and Hinkley, D. V. (1974), {\em Theoretical Statistics},
London: Chapman \& Hall.

\bibitem{ci80}
Cox, D. R., and Isham, V. (1980), {\em Point Processes},
London: Chapman \&  Hall.

\bibitem{cl66}
Cox, D. R., and Lewis, P. A. W. (1966), {\em The Statistical Analysis
of a Series of Events}, London: Chapman \& Hall.

\bibitem{co84} 
Cox, D. R., and Oakes, D. (1984), {\em Analysis of Survival Data},
London: Chapman \& Hall.

\bibitem{cs68}
Cox, D. R., and Snell, E.\ J., (1968), A general definition of
residuals (with discussion) {\em Journal of the Royal Statistical
Society B}, {\bf 30}, 248--275.


\bibitem{cs94}
Cragnolino, G., and Sridhar, N., Editors (1994), {\em Application of
Accelerated Corrosion Tests to Service Life Prediction of Materials},
ASTM STP 1194, Philadelphia: American Society for Testing and Materials.

\bibitem{cs88}
Crow, E. L., and Shimizu, K. (1988), {\em Lognormal Distributions:
Theory and Applications}, New York: Marcel Dekker.

\bibitem{crow82}
Crow, L. (1982), Confidence interval procedures for the Weibull
process with applications to reliability growth, {\em Technometrics},
{\bf 24}, 67--72.

\bibitem{ckss91}
Crowder, M. J., Kimber, A. C., Smith, R. L., and Sweeting, T. J.
(1991), {\em Statistical Analysis of Reliability Data}, New York:
Chapman \& Hall.

\bibitem{crowder97}
Crowder, S. V. (1997), How to determine component-based preventive
maintenance plans, SEMATECH Technology Transfer Document
\#92051125A-GEN, SEMATECH, Austin, TX.

\bibitem{dale85}
Dale, C. J. (1985), Application of the proportional hazards model in
the reliability field, {\em Reliability Engineering}, {\bf 10}, 5--25.

\bibitem{david81}
David, H. A. (1981), {\em Order Statistics} (Second Edition), New
York: John Wiley \& Sons.

\bibitem{dm78}
David, H. A., and Moeschberger, M. L. (1978), {\em The Theory of Competing
Risks}, London: Griffin.

\bibitem{dg95}
Davidian, M., and Giltinan, D. M. (1995), {\em Nonlinear Models for
Repeated Measurement Data}, London: Chapman \&  Hall.

\bibitem{davis52}
Davis, D. J. (1952), An analysis of some failure data, {\em Journal of
the American Statistical Association}, {\bf 47}, 113--150.

\bibitem{day69}
Day, N. E. (1969), Estimating the components of a mixture of normal
distribution, {\em Biometrika, } {\bf 56}, 463--474.

\bibitem{deming75}
Deming, W. E. (1975), On probability as a basis for action, {\em The
American Statistician}, {\bf 29}, 146--152.

\bibitem{derringer82}
Derringer, G. C. (1982), Considerations in single and multiple
stress accelerated life testing, {\em Journal of Quality
Technology}, {\bf 14}, 130--134.

\bibitem{desmond86}
Desmond, A. F. (1986), On the relationship between two fatigue-life
models, {\em IEEE Transactions on Reliability}, {\bf R-35}, 167--169.

\bibitem{doganaksoy95}
Doganaksoy, N.  (1995), Likelihood ratio confidence intervals in
life-data analysis, Chapter 20 in {\em Recent Advances in
Life-Testing and Reliability}, N. Balakrishnan, Editor.  Boca Raton,
FL: CRC Press.

\bibitem{dn91}
Doganaksoy, N., and Nelson, W. (1991), A method and computer program
MCFDIFF to compare two samples of repair data, TIS report 91CRD172,
General Electric Company Research and Development, Schenectady, NY.

\bibitem{ds93} 
Doganaksoy, N.,  and Schmee, J. (1993), Comparisons of approximate
confidence intervals for distributions used in life-data analysis,
{\em Technometrics}, {\bf 35}, 175--184.

\bibitem{Dowing93}
Dowling, N. E. (1993), {\em Mechanical Behavior of Materials},
Englewood Cliffs, NJ: Prentice Hall.

\bibitem{Drapella}
Drapella, A. (1992), An extended mathematical model for failure
kinetics, {\em Quality and Reliability Engineering International},
{\bf 8}, 371--373.

\bibitem{drenick60}
Drenick, R. F. (1960), The failure law of complex equipment, {\em
Journal of the Society of Industrial and Applied Mathematics}, {\bf
8}, 680--690.

\bibitem{efron81}
Efron, B. (1981), Censored data and the bootstrap, {\em Journal of the
American Statistical Association}, {\bf 76}, 312--319.

\bibitem{efron92}
Efron, B. (1982), {\em The Jackknife, the Bootstrap, and Other
Resampling Plans}, Philadelphia: Society for Industrial and Applied
Mathematics.

\bibitem{efron85}
Efron, B.\ (1985), Bootstrap confidence intervals for a class of
parametric problems, {\em Biometrika}, {\bf 72}, 45---58.

\bibitem{et93}
Efron, B., and Tibshirani, R. J. (1993), {\em An Introduction to the
Bootstrap}, New York: Chapman \& Hall.

\bibitem{ejj80}
Elandt-Johnson, R. C., and Johnson, N. L. (1980), {\em Survival Models
and Data Analysis}, New York: John Wiley \& Sons.

\bibitem{eb79}
Engelhardt, M., and Bain, L. J. (1979),
Prediction limits and two-sample problems with complete or
censored Weibull data, {\em Technometrics, } {\bf 21, }233--237.

\bibitem{ebw81}
Engelhardt, M., Bain, L. J., and Wright, F. T. (1981), Inferences on
the parameters of the Birnbaum-Saunders fatigue life distribution
based on maximum likelihood estimation, {\em Technometrics}, {\bf 23},
251--256.

\bibitem{es53}
Epstein, B., and Sobel, M. (1953), Life testing,  {\em Journal of
the American Statistical Association}, {\bf 48}, 486--502.

\bibitem{em92}
Escobar, L. A., and Meeker, W. Q. (1992), Assessing influence in
regression analysis with censored data, {\em Biometrics}, {\bf 48},
507--528.

\bibitem{em94}
Escobar, L. A., and Meeker, W. Q. (1994), Fisher information matrix
for the extreme value, normal, and logistic distributions and censored
data, {\em Applied Statistics}, {\bf 43}, 533--540.


\bibitem{em95}
Escobar, L. A., and Meeker, W. Q. (1995), Planning accelerated life
tests with two or more experimental factors, {\em Technometrics}, {\bf 37},
411--427.

\bibitem{em97b}
Escobar, L. A., and Meeker, W. Q. (1998a), Statistical prediction
based on censored life data, Department of Statistics,
Iowa State University, Ames, IA.

\bibitem{em97c}
Escobar, L. A., and Meeker, W. Q. (1998b), Simultaneous confidence
bands for models based on location-scale distributions and censored
data, Department of Statistics,
Iowa State University, Ames, IA.

\bibitem{em97d}
Escobar, L. A., and Meeker, W. Q. (1998c), Stable parameterizations
for maximum likelihood methods with censored data, Department of
Statistics, Iowa State University, Ames, IA.

\bibitem {em98}
Escobar, L. A., and Meeker, W. Q. (1998d), Fisher information matrices
with censoring, truncation, and explanatory variables, {\em Statistica
Sinica}, {\bf 7}, xxx--xxx (in press).

\bibitem{ehp93}
Evans, M., Hastings, N., and Peacock, B.  (1993), {\em Statistical
Distributions} (Second Edition), New York: Wiley-Interscience.

\bibitem{evans77}
Evans, R. A. (1977), Accelerated testing: the only game in town,
{\em IEEE Transactions on Reliability}, {\bf R-26}, 241.

\bibitem{evans89}
%october 89 issue
Evans, R. A. (1989), Bayes is for the birds, {\em IEEE Transactions on
Reliability}, {\bf R-38}, 401.

\bibitem{evans91}
Evans, R. A. (1991), Accelerated testing, {\em IEEE Transactions on
Reliability}, {\bf R-40}, 497.

\bibitem{eh81}
Everitt, B. S., and Hand, D. J. (1981), {\em Finite Mixture
Distributions}, London: Chapman \& Hall.

\bibitem{eyring80}
Eyring, H. (1980), {\em Basic Chemical Kinetics}, New York: John Wiley \& Sons.

\bibitem{egl41}
Eyring, H., Gladstones, S., and Laidler, K. J. (1941), {\em The Theory
of Rate Processes}, New York:
McGraw Hill.

\bibitem{falls70}
Falls, L. W. (1970), Estimation of parameters in compound Weibull
distributions, {\em Technometrics}, {\bf 12}, 399--407.

\bibitem{fc79}
Farewell, V. T., and Cox, D. R. (1979), A note on multiple time scales
in life testing, {\em Applied Statistics}, {\bf 28}, 73--75.

\bibitem{fp77}
Farewell, V. T., and Prentice, R. L. (1977), A study of distributional
shape in life testing, {\em Technometrics}, {\bf 19}, 69--75.

\bibitem{faulkenberry73}
Faulkenberry, G. D. (1973), A method of obtaining prediction intervals,
{\em Journal of the American Statistical Association}, 
{\bf 68}, 433--435.

\bibitem{fw95}
Feinberg, A. A., and Windom, A. (1995), The reliability physics of
thermodynamic aging, Chapter 13 in {\em Recent Advances in
Life-Testing and Reliability}, N. Balakrishnan, Editor. Boca Raton,
FL: CRC Press.

\bibitem{fisher25}
Fisher, R. A. (1925), Theory of statistical estimation, {\em
Proceedings of the Cambridge Philosophical Society}, {\bf 22}, 700--725.

\bibitem{fh92}
Fleming, T. R., and Harrington, D. P. (1992), {\em Counting Processes
and Survival Analysis}, New York: John Wiley \& Sons.

\bibitem{fg80}
Friedman, L. B., and Gertsbakh, I. (1980), Maximum likelihood
estimation in a minimum-type model with exponential and Weibull
failure modes, {\em Journal of the American Statistical Association},
{\bf 75}, 460--465.

\bibitem{fukuda91}
Fukuda, M. (1991), {\em Reliability and Degradation of Semiconductor
Lasers and LED's}, Norwood, MA: Artech House.

\bibitem{galambos78}
Galambos, J. (1978), {\em The Asymptotic Theory of Extreme Order
Statistics}, New York: John Wiley \& Sons.

\bibitem{geisser93}
Geisser, S. (1993), {\em Predictive Inference: An Introduction}, New
York: Chapman \& Hall.

\bibitem{gs90}
Gelfand, A. E., and Smith, A. F. M. (1990), Sampling-based approach to
calculating marginal densities, {\em Journal of the American
Statistical Association}, {\bf 85}, 398--409.

\bibitem{gs92}
Gelfand, A. E., and Smith, A. F. M. (1992), Bayesian analysis of
constrained parameter and truncated data problems using Gibbs
sampling, {\em Journal of the American Statistical Association}, {\bf 87},
523--532.

\bibitem{gcsr95}
Gelman, A., Carlin, J. B., Stern, H. S., and Rubin, D. B. (1995),
{\em Bayesian Data Analysis},
New York: Chapman \& Hall.

\bibitem{gg93}
Gentleman, R., and Geyer, C. J. (1994), Maximum likelihood for
interval censored data: consistency and computation, {\em Biometrika},
{\bf 81}, 618--623.

\bibitem{gertsbakh89}
Gertsbakh, I. B. (1989), {\em Statistical Reliability Theory}, New
York: Marcel Dekker.

\bibitem{gk69}
Gertsbakh, I. B., and Kordonsky, Kh. B. (1969), {\em Models of
Failure}, English translation from the Russian version, New York:
Springer-Verlag.

\bibitem{gk76} 
Giesbrecht, F., and Kempthorne, O. (1976), Maximum likelihood
estimation in the three-parameter lognormal distribution,  {\em
Journal of the Royal Statistical Society B}, {\bf 38}, 257--264.

\bibitem{gc85}
Gillen, K. T., and Clough, R. L. (1985), A kinetic model for predicting
oxidative degradation rates in combined radiation-thermal
environments, {\em Journal of Polymer Science, Polymer Chemistry
Edition}, {\bf 23}, 2683--2707.

\bibitem{gm85}
Gillen, K. T., and Mead, K. E. (1980), Predicting life expectancy
and simulating age of complex equipment using accelerated aging
techniques.  Available from the National Technical Information
Service, U. S. Department of Commerce, 5285 Port Royal Road,
Springfield, VA 22151.

\bibitem{gu95}
Gnedenko, B., and Ushakov, I. (1995), {\em Probabilistic Reliability
Engineering}, New York: John Wiley \& Sons.

\bibitem{gm96}
Graves, S., and Menten, T. (1996), Designing experiments to measure
and improve reliability, Chapter 11 in {\em Handbook of Reliability
Engineering and Management} (Second Edition), W. G. Ireson, C. F
Coombs, and R. Y.  Moss, Editors.  New York: McGraw Hill.

\bibitem{griffiths80}
Griffiths, D. A.  (1980), Interval estimation for the three-parameter
lognormal distribution via the likelihood function, {\em Applied
Statistics}, {\bf 29}, 58--68.

\bibitem{guh91}
Guess, F. M., Usher, J. S., and Hodgson, T. J. (1991), Estimating
system and component reliabilities under partial information on cause
of failure, {\em Journal of Statistical Planning and Inference},
{\bf 29}, 75--85.


\bibitem{hm82a}
Hahn, G. J., and Meeker, W. Q. (1982a), Pitfalls and practical
considerations in product life analysis, part 1: basic concepts and
dangers of extrapolation,  {\em Journal of Quality Technology},
{\bf 14}, 144--152.

\bibitem{hm82b}
Hahn, G. J., and Meeker, W. Q. (1982b), Pitfalls and practical
considerations in product life analysis, part 2: mixtures of product
populations and more general models, {\em Journal of Quality
Technology}, {\bf 14}, 177--185.

\bibitem{hm91}
Hahn, G. J., and Meeker, W. Q. (1991), {\em Statistical Intervals: A
Guide for Practitioners}, New York: John Wiley \& Sons.

\bibitem{hn93}
Hahn, G. J., and Nelson, W. (1973), A survey of prediction intervals
and their applications, {\em Journal of Quality Technology, } {\bf 5},
178--188.

\bibitem{hs67}
Hahn, G. J., and Shapiro, S. S. (1967), {\em Statistical Models in
Engineering}, New York: John Wiley \& Sons.

\bibitem{hakim89}
Hakim, E. B. (1989), {\em Microelectronic Reliability, Volume I,
Reliability, Test and Diagnostics}, Norwood, MA: Artech House.

\bibitem{hall92}
Hall, P. (1992), {\em The Bootstrap and Edgeworth Expansion}, New
York: Springer-Verlag.

\bibitem{Hamada93}
Hamada, M. (1993), Reliability improvement via Taguchi's robust
design, {\em Quality and Reliability Engineering International},
{\bf 9}, 7--13.

\bibitem{Hamada95a}
Hamada, M. (1995a), Using statistically designed experiments to
improve reliability and to achieve robust reliability, {\em IEEE
Transactions on Reliability}, {\bf R-44}, 206--215.

\bibitem{hamada95}
Hamada, M. (1995b), Analysis of experiments for reliability
improvement and robust reliability, Chapter 9 in {\em Recent
Advances in Life-Testing and Reliability}, N. Balakrishnan,
Editor. Boca Raton, FL: CRC Press.


\bibitem{hw92b} 
Hamada, M., and Wu, C. F. J. (1995), Analysis of censored data from
fractionated experiments: a Bayesian approach, {\em Journal of the
American Statistical Association}, {\bf 90}, 467--477.

\bibitem{harter77}
Harter, H. L. (1977), A survey of the literature on the size effect
of material strength, Report No. AFFDL-TR-77-11, Air Force Flight
Dynamics Lab AFSC, Wright-Patterson AFB, OH 45433.

\bibitem{harter84}
Harter, H. L. (1984), Another look at plotting positions, {\em
Communications in Statistics, Part B---Simulation and Computation},
{\bf 13}, 1613--1633.

\bibitem{hl97}
Hirose, H., and Lai, T. L. (1997), Inference from grouped data in
three-parameter Weibull models with applications to breakdown voltage
experiments, {\em Technometrics}, {\bf 39}, 199--210.

\bibitem{ha90}
Hooper, J. H.,  and Amster, S. J. (1990), Analysis and presentation
of reliability data, in {\em Handbook of Statistical Methods for
Engineers and Scientists}, Harrison M. Wadsworth, Editor. New York:
McGraw Hill.

\bibitem{hosmer73}
Hosmer, D. W. Jr. (1973), A comparison of iterative maximum
likelihood estimates of the parameters of a mixture of two normal
distributions under three different types of sample, {\em Biometrics},
{\bf 29}, 761--770.

\bibitem{hm81}
Howes, M. J., and Morgan, D. V. (1981), {\em Reliability and
Degradation}, New York: John Wiley \&
Sons.

\bibitem{hr94}
H\o yland, A., and Rausand, M. (1994), {\em System Reliability Theory:
Models and Statistics Methods}, New York: John Wiley \& Sons.

\bibitem{hsbm78}
Hudak, S. J. Jr.,  Saxena, A., Bucci, R. J., and Malcolm, R. C.\ (1978),
Development of standard methods of testing and analyzing fatigue
crack growth rate data, Technical Report AFML-TR-78-40, Westinghouse
R \& D Center, Westinghouse Electric Corporation, Pittsburgh,
PA 15235.

\bibitem{ireson96}
Ireson, W. G. (1996), Reliability information collection and analysis,
Chapter 10 in {\em Handbook of Reliability Engineering and
Management} (Second Edition), W. G. Ireson, C. F. Coombs, and R. Y.
Moss, Editors.  New York: McGraw Hill.


\bibitem{jm97}
Jeng, S. L., and Meeker, W. Q. (1998), Comparisons of Weibull
distribution confidence intervals for type I censored data, Department
of Statistics, Iowa State University, Ames, IA.


\bibitem{jm90}
Jensen, K. L., and Meeker, W. Q. (1990), ALTPLAN: microcomputer
software for developing and evaluating accelerated life test plans,
paper presented at the Annual ASA Meetings, Anaheim, CA.

\bibitem{jensen95}
Jensen, F. (1995), {\em Electronic Component Reliability:
Fundamentals, Modelling, Evaluation, and Assurance}, New York: John
Wiley \& Sons.

\bibitem{jp82}
Jensen, F., and Petersen, N.~E. (1982), {\em Burn-in: An Engineering
Approach to Design and Analysis of Burn-in Procedures}, New York: John
Wiley \& Sons.

\bibitem{jkb94}
Johnson, N. L., Kotz, S., and Balakrishnan, N. (1994), {\em
Continuous Univariate Distributions Volume 1}, New York: John Wiley
\& Sons.

\bibitem{jkb95}
Johnson, N. L., Kotz, S., and Balakrishnan, N.  (1995), {\em
Continuous Univariate Distributions Volume 2}, New York: John Wiley
\& Sons.

\bibitem{jlnbh85}
Joyce, W. B., Liou, K-Y., Nash, F. R., Bossard, P. R., and Hartman,
R.  L. (1985), Methodology of accelerated aging, {\em AT\&T
Technical Journal}, {\bf 64}, 717--764.

\bibitem{kalb71}
Kalbfleisch, J. D. (1971), Likelihood methods of prediction,  in {\em
Foundations of Statistical Inference; proceedings of the Symposium on
the Foundations of Statistical Inference}, V. P. Godambe,  and
 D. A. Sprott, Editors. Toronto: Holt, Rinehart and Winston of Canada.

\bibitem{kl88}
Kalbfleisch, J. D., and Lawless, J. F.  (1988),
Estimation of reliability in field-performance studies,
{\em Technometrics}, {\bf 30}, 365--388 (with discussion).

\bibitem{kl92}
Kalbfleisch, J. D., and Lawless, J. F. (1992), Some useful statistical
methods for truncated data, {\em Journal of Quality Technology}, {\bf
24}, 145--152.

\bibitem{kp80}
Kalbfleisch, J. D., and Prentice, R. L. (1980), {\em The Statistical
Analysis of Failure Time Data}, New York: John Wiley \& Sons.

\bibitem{kr89}
Kalkanis, G., and Rosso, E. (1989),
The inverse power law model for the lifetime of a mylar--polyurethane
laminated DC HV insulating structure, {\em Nuclear Instruments and
Methods in Physics Research}, {\bf A281}, 489--496.

\bibitem{km58} 
Kaplan, E. L., and Meier, P. (1958), Nonparametric estimation from
incomplete observations, {\em Journal of the American Statistical
Association}, {\bf 53}, 457--481.

\bibitem{ks95}
Kececioglu, D., and Sun, F. (1995), {\em Environmental Stress Screening: Its
Quantification, Optimization and Management}, Englewood Cliffs, NJ:
 Prentice Hall.

\bibitem{kf71}
Kempthorne, O., and Folks, L. (1971), {\em Probability, Statistics, and
Data Analysis}, Ames, IA: Iowa State University Press.

\bibitem{kg80}
Kennedy, W. J., and Gentle, J. E. (1980), {\em Statistical Computing, } New
York: Marcel Dekker.

\bibitem{klinger91a}
Klinger, D. J. (1991a), On the notion of activation energy in
reliability: Arrhenius, Eyring, and thermodynamics,  {\em 1991
Proceedings of the Annual Reliability and Maintainability
Symposium}, 295--300, New York: Institute of
Electrical and Electronics Engineers.

\bibitem{klinger91b}
Klinger, D. J. (1991b), Humidity acceleration factor for plastic
packaged electronic devices, {\em Quality and Reliability
Engineering International}, {\bf 7}, 365--370.

\bibitem{klinger92}
Klinger, D. J. (1992), Failure time and rate constant of degradation:
an argument for the inverse relationship, {\em Microelectronics and
Reliability}, {\bf 32}, 987--994.

\bibitem{knm90}
Klinger, D. J., Nakada, Y., and Menendez, M. A. (1990),
{\em AT\&T Reliability Manual}, New York: Van Nostrand Reinhold.

\bibitem{klion92}
Klion, J. (1992), {\em Practical Electronic Reliability Engineering},
New York: Van Nostrand Reinhold.

\bibitem{knezevic93}
Knezevic, J. (1993), {\em Reliability, Maintainability and
Supportability}, London: McGraw Hill.

\bibitem{kg93}
Kordonsky, Kh. B., and Gertsbakh, I. B. (1993), Choice of the best
time scale for system reliability analysis, {\em European Journal of
Operational Research}, {\bf 65}, 235--246.

\bibitem{kg95a}
Kordonsky, Kh. B., and Gertsbakh, I. B. (1995a), System state
monitoring and lifetime scales---I, {\em Reliability Engineering and
System Safety}, {\bf 47}, 1--14.

\bibitem{kg95b}
Kordonsky, Kh. B., and Gertsbakh, I. B. (1995b), System state
monitoring and lifetime scales---II, {\em Reliability Engineering and
System Safety}, {\bf 49}, 145--154.

\bibitem{ku70}
Kozlov, B. A., and Ushakov, I. A. (1970), {\em Reliability
Handbook}, New York: Holt, Rinehart, and Winston.

\bibitem{kull61}
Kulldorff, G. (1961), {\em Contributions to the Theory of Estimation
from Grouped and Partially Grouped Samples}, Stockholm: Almqvist and
Wiksell.

\bibitem{kuo84}
Kuo, W. (1984), Reliability enhancement through optimal burn-in,
{\em IEEE Transactions on Reliability}, {\bf R-33}, 145--156.


\bibitem{kck98}
Kuo, W., Chien, W. T. K., and Kim, T. (1998),
{\em Reliability and Stress Burn-in},
Netherlands: Kluwer Academic Publishers.

\bibitem{ky96}
Kuo, L., and Yang, T. L. (1996), Bayesian computation for
nonhomogeneous Poisson processes in software reliability, {\em
Journal of the American Statistical Association}, {\bf 91}, 763--773.

\bibitem{lagakos79}
Lagakos, S. W. (1979), General right censoring and its impact on the
analysis of survival data, {\em Biometrics}, {\bf 35}, 139--156.

\bibitem{lk97}
Landers, T. L., and Kolarik, W. J. (1987),
Proportional hazards analysis of field warranty data,
{\em Reliability Engineering}, {\bf 18}, 131--139.

\bibitem{lawless73}
Lawless, J. F. (1973), On estimation of safe life when the underlying
distribution is Weibull, {\em Technometrics}, {\bf 15}, 857--865.

\bibitem{lawless82}
Lawless, J. F. (1982), {\em Statistical Models and Methods for
Lifetime Data}, New York: John Wiley \& Sons.

\bibitem{lawless86}
Lawless, J. F. (1986), A note of lifetime regression models, {\em
Biometrika}, {\bf 73}, 509--512.

\bibitem{lawless98}
Lawless, J. F. (1998), Statistical analysis of product warranty
data, {\em International Statistical Review}, {\bf xxx}, xxx--xxx (in
press).

\bibitem{lhc95}
Lawless, J. F., Hu, J., and Cao, J. (1995), Methods for the
estimation of failure distributions and rates from automobile
warranty data, {\em Lifetime Data Analysis}, {\bf 1}, 227--240.

\bibitem{lk92}
Lawless, J. F., and Kalbfleisch, J. D. (1992), Some issues in the
collection and analysis of field reliability data, in {\em Survival
Analysis: State of the Art}, 141--152, J. P. Klein and P. K. Goel,
Editors. Netherlands: Kluwer Academic Publishers.

\bibitem{ln95}
Lawless, J. F., and Nadeau, C. (1995), Some simple robust methods for
the analysis of recurrent events, {\em Technometrics}, {\bf 37},
158--168.

\bibitem{lt96}
Lawless, J. F., and Thiagarajah, K. (1996), A point-process
model incorporating renewals and time trends, with
application to repairable systems, {\em
Technometrics}, {\bf 38}, 131--138.

\bibitem{lecam90} 
Le Cam, L. (1990), Maximum likelihood: an introduction, {\em
International Statistical Review}, {\bf 58}, 153--171.

\bibitem{lee80}
Lee, L.  (1980), Testing adequacy of the Weibull and loglinear rate 
models for a Poisson process, {\em
Technometrics}, {\bf 22}, 195--199.

\bibitem{lehmann83}
Lehmann, E. L. (1983), {\em Theory of Point Estimation}, New York:
John Wiley \& Sons.

\bibitem{lm97}
Lerch, C. F., and Meeker, W. Q. (1998), A Bayes approach to analyzing
accelerated life test data, Department of Statistics, Iowa State
University, Ames,  IA.

\bibitem{lewis96}
Lewis, E. E. (1996), {\em Introduction to Reliability Engineering},
New York: John Wiley \& Sons.

\bibitem{li95a}
Li, G. (1995a), On nonparametric likelihood ratio estimation of
survival probabilities for censored data, {\em Statistics \&
Probability Letters}, {\bf 25}, 95--104.

\bibitem{li95b}
Li, G. (1995b), Nonparametric likelihood ratio estimation of
probabilities for truncated data, {\em Journal of the American
Statistical Association}, {\bf 90}, 997--1003.

\bibitem{lz56}
Lieblein J., and Zelen, M. (1956), Statistical investigation of the
fatigue life of deep-groove ball bearings, {\em Journal of Research,
National Bureau of Standards}, {\bf 57}, 273--316.

\bibitem{lindley72}
Lindley, D. V. (1972), {\em Bayesian Statistics, A Review},
Philadelphia: Society for Industrial and Applied Mathematics.

\bibitem{lindsey96}
Lindsey, J. K. (1996), {\em Parametric Statistical Inference},
Oxford: Clarendon Press.

\bibitem{lb90}
Lindstrom, M. J., and Bates, D. M. (1990), Nonlinear mixed effects
models for repeated measures data, {\em Biometrics}, {\bf 46},
673--687.

\bibitem{liume97}
Liu, S., Meeker, W. Q., and Escobar, L. A. (1998), A stable
parameterization for fitting the generalized gamma distribution to
censored data,  Department of Statistics, Iowa State University,
Ames, IA.

\bibitem{lp89}
Lu, J.\ C., and Pantula, S.\ G.\ (1989), A repeated-measurements
model for over-stressed degradation data, Department of Statistics,
North Carolina State University, Raleigh, NC  27695-8203.

\bibitem{lm93}
Lu, C. J., and Meeker, W. Q. (1993), Using degradation measures to
estimate a time-to-failure distribution, {\em Technometrics}, {\bf
34}, 161--174.

\bibitem{lme96}
Lu, C. J., Meeker, W. Q., and Escobar, L. A. (1996), A comparison of
degradation and failure-time analysis methods of estimating a
time-to-failure distribution,  {\em Statistica Sinica}, {\bf 6},
531--546.

\bibitem{luvalle90}
LuValle, M. J. (1990), A note on experiment design for accelerated
life tests, {\em Microelectronics and Reliability}, {\bf 30}, 591--603.

\bibitem{luvalle93}
LuValle, M. J. (1993), Experimental design and graphical analysis
for checking acceleration models, {\em Microelectronics and
Reliability}, {\bf 33}, 741--763.

\bibitem{lh92}
LuValle, M. J., and Hines, L. L. (1992), Using step stress to
explore the kinetics of failure, {\em Quality and Reliability
Engineering International}, {\bf 8}, 361--369.

\bibitem{lwm86}
LuValle, M. J., Welsher, T. L., and Mitchell, J. P. (1986),
A new approach to the extrapolation of accelerated
life test data, {\em The Proceedings of the Fifth International
Conference on Reliability and Maintainability}, 620--635,
Biarritz, France.

\bibitem{lws88}
LuValle, M. J., Welsher, T. L., and Svoboda, K. (1988),
Acceleration transforms and statistical kinetic models,
{\em Journal of Statistical Physics}, {\bf 52}, 311--320.

\bibitem{mss74}
Mann, N. R., Schafer, R. E., and Singpurwalla, N. D. (1974), {\em
Methods for Statistical Analysis of Reliability and Life Data}, New
York: John Wiley \& Sons.

\bibitem{martin82}
Martin, J. W. (1982), Time transformation functions commonly used in
life testing analysis, {\em Durability of Building Materials}, {\bf
1}, 175--194.

\bibitem{mw82}
Martz, H. F., and Waller, R. A. (1982), {\em Bayesian Reliability
Analysis}, New York: John Wiley \& Sons.

\bibitem{mccool80}
McCool, J. I. (1980), Confidence limits for Weibull regression with
censored data, {\em IEEE Transactions on Reliability}, {\bf R-29},
145--150.

\bibitem{mb85}
McPherson, J. W.,  and Baglee, D. A. (1985), Acceleration factors for
thin gate oxide stressing, {\em Proceedings of the IEEE International
Reliability Physics Symposium}, {\bf 23}, 1--5.

\bibitem{mk93}
Mee, R., and Kushary, D. (1994), Prediction limits for the Weibull
distribution utilizing simulation, {\em Computational Statistics and
Data Analysis}, {\bf 17}, 327--336.

\bibitem{meeker84}
Meeker, W. Q. (1984),
A comparison of accelerated life test plans for
Weibull and lognormal distributions and Type I censored data,
{\em Technometrics}, {\bf 26}, 157--171.

\bibitem{meeker86} 
Meeker, W. Q. (1986), Planning life tests in which units are inspected
for failure, {\em IEEE Transactions on Reliability}, {\bf R-35},
571--578.

\bibitem{Meeker87} 
Meeker, W. Q.  (1987), Limited failure population life tests:
application to integrated circuit reliability,  {\em Technometrics},
{\bf 29}, 51--65.

\bibitem {me93}
Meeker, W. Q., and Escobar, L. A. (1993), A review of recent research
and current issues in accelerated testing,   {\em International
Statistical Review}, {\bf 61}, 147--168.

\bibitem  {me94}
Meeker, W. Q., and Escobar, L. A. (1994), Maximum likelihood methods
for fitting parametric statistical models to censored and truncated
data, in {\em Probabilistic and Statistical Methods in the
Physical Sciences},  J. Stanford and S. Vardeman, Editors.
New York: Academic Press.

\bibitem{me95}
Meeker, W. Q., and Escobar, L. A. (1995), Teaching about
approximate confidence regions based on maximum likelihood estimation,
{\em The American Statistician}, {\bf 49}, 48--53.

\bibitem {meh92}
Meeker, W. Q., Escobar, L. A., and Hill, D. A.  (1992), Sample sizes for
estimating the Weibull hazard function from censored samples, {\em
IEEE Transactions on Reliability}, {\bf 41}, 133--138.

\bibitem{mel96}
Meeker, W. Q., Escobar, L. A., and Lu, C. J. (1998), Accelerated
degradation tests: modeling and analysis,  {\em Technometrics},
{\bf 40}, xxx--xxx.

\bibitem{mh85}
Meeker, W. Q., and Hahn, G. J. (1985), {\em How to Plan Accelerated
Life Tests: Some Practical Guidelines}, Volume 10 of the ASQC Basic
References in Quality Control: Statistical Techniques. Available from
the American Society for Quality Control, 310 W. Wisconsin Ave.,
Milwaukee, WI 53203.

\bibitem{mh95}
Meeker, W. Q., and Hamada, M. (1995), Statistical tools for the
rapid development \& evaluation of high-reliability products, {\em
IEEE Transactions on Reliability}, {\bf R-44}, 187--198.

\bibitem{ml95} 
Meeker, W. Q., and LuValle, M. J. (1995), An accelerated life test
model based on reliability kinetics, {\em Technometrics},
{\bf 37}, 133--146.

\bibitem{mn75}
Meeker, W. Q., and Nelson, N. (1975), Optimum accelerated life tests
for Weibull and extreme value distributions, {\em IEEE Transactions
on Reliability}, {\bf R-24}, 321--332.

\bibitem{mn76} 
Meeker, W. Q., and Nelson, W. (1976), Weibull percentile estimates and
confidence limits from singly censored data by maximum likelihood,
{\em IEEE Transactions on Reliability}, {\bf R-25}, 20--24.

\bibitem{mn77} 
Meeker, W. Q., and Nelson, W. (1977), Weibull variances 
and confidence limits by maximum likelihood
for singly censored data, {\em Technometrics}, {\bf 19},
473--476.

\bibitem{mm94} 
Meeter, C. A., and Meeker, W. Q. (1994), Optimum accelerated life
tests with a nonconstant scale parameter, {\em Technometrics}, {\bf
36}, 71--83.

\bibitem{mil81} 
MIL-HDBK-189 (1981), {\em Reliability Growth Management}.
Available from Naval Publications and Forms Center, 5801 Tabor Ave,
Philadelphia, PA 19120.

\bibitem{milhdbk217e} 
MIL-HDBK-217E (1986), {\em Reliability Prediction for Electronic Equipment}.
Available from Naval Publications and Forms Center, 5801 Tabor Ave,
Philadelphia, PA 19120.

\bibitem{mil85} 
MIL-STD-883 (1985), {\em Test Methods and Procedures for Microelectronics}.
Available from Naval Publications and Forms Center, 5801 Tabor Ave,
Philadelphia, PA 19120.

\bibitem{milstd1629} 
MIL-STD-1629A (1980), {\em Failure Modes and Effects Analysis}.
Available from Naval Publications and Forms Center, 5801 Tabor Ave,
Philadelphia, PA 19120.

\bibitem{milstd2164} 
MIL-STD-2164 (1985),  {\em Environmental Stress Screening Process for
Electronic Equipment}.
Available from Naval Publications and Forms Center, 5801 Tabor Ave,
Philadelphia, PA 19120.

\bibitem{minitab97} 
MINITAB (1997), Minitab User's Guide 2: Data Analysis and
Quality Tools, Release 12, Minitab., State College, PA.

\bibitem{morgan84} 
Morgan, B. J. T. (1984), {\em Elements of Simulation}, 
London: Chapman \& Hall.

\bibitem{morgan80} 
Morgan, C. B. (1980), Analyzing competing failure modes using statpac
and a simple actuarial technique, General Electric Corporate Research
and Development technical report 80CRD110, Schenectady, NY.

\bibitem{mm98}
Morse, P. N., and Meeker, W. Q. (1998), Simulation methods for planning
an accelerated degradation test, Department of Statistics, Iowa State
University, Ames, IA.

\bibitem{murray93}
Murray, W. P. (1993), Archival life expectancy of 3M magneto-optic
media,  {\em Journal of the Magnetics Society of Japan}, {\bf
17}, Supplement S1, 309--314.


\bibitem{murray94}
Murray, W. P. (1994), Accelerated service life prediction of compact
disks, in {\em Accelerated and Outdoor Durability Testing of Organic
Materials, ASTM STP 1202}, 263--271. W. D. Ketola and D. Grossman,
Editors. Philadelphia: American Society for Testing and Materials.

\bibitem{mm96}
Murray, W. P., and Maekawa, K. (1996), Reliability evaluation of 3M
magneto-optic media, {\em Journal of the Magnetics Society of
Japan}, {\bf 20}, Supplement S1, 309--314.


\bibitem{mio87}
Musa, J. D., Iannino, A., and Okumoto, K.  (1987), {\em Software
Reliability: Measurement, Prediction, Application}, New York:
McGraw Hill.

\bibitem{nair81}
Nair, V. N. (1981), Plots and tests for goodness of fit with randomly
censored data, {\em Biometrika }, {\bf 68}, 99--103.

\bibitem{nair84}
Nair, V. N. (1984), Confidence bands for survival functions with
censored data: a comparative study, {\em Technometrics}, {\bf 26},
265--275.

\bibitem{nakamura91}
Nakamura, T. (1991), Existence of maximum likelihood estimates for
interval-censored data from some three-parameter models with a shifted
origin, {\em Journal of the Royal Statistical Society B}, {\bf 53},
211--220.

\bibitem{nq96}
Nash, J. C., and Quon, T. K. (1996), Software for modeling kinetic
phenomena, {\em The American Statistician}, {\bf 50}, 368--378.

\bibitem{nelson69}
Nelson, W. (1969), Hazard plotting for incomplete failure data, {\em
Journal of Quality Technology}, {\bf 1}, 27--52.

\bibitem{nelson72}
Nelson, W. (1972), Theory and application of hazard plotting for
censored survival data, {\em Technometrics}, {\bf 14}, 945--966.

\bibitem{nelson73}
Nelson, W. (1973), Analysis of residuals from censored data,
{\em Technometrics}, {\bf 15}, 697--715.

\bibitem{nelson75}
Nelson, W. (1975a), Analysis of accelerated life data with a mix of
failure modes by maximum likelihood,
{\em IEEE Transactions on Reliability}, {\bf R-24}, 230--237.

\bibitem{nelson75a}
Nelson, W. (1975b), Graphical analysis of accelerated life data with a
mix of failure modes, {\em IEEE Transactions on Reliability}, {\bf
R-24}, 230--237.

\bibitem{nelson81}
Nelson, W. (1981), Analysis of performance degradation data from
accelerated tests, {\em IEEE Transactions on Reliability}, {\bf
R-30}, 3, 149--155.

\bibitem{nelson82}
Nelson, W. (1982), {\em Applied Life Data Analysis}, New York: John
Wiley \& Sons.

\bibitem{nelson84}
Nelson, W. (1984), Fitting of fatigue curves with nonconstant standard
deviation to data with runouts, {\em Journal of Testing and Evaluation
}, {\bf 12}, 69--77.

\bibitem{nelson85}
Nelson, W. (1985), Weibull analysis of reliability data with few or no
failures, {\em Journal of Quality Technology}, {\bf 17}, 140--146.

\bibitem{nelson88}
Nelson, W. (1988), Graphical analysis of system repair data,
{\em Journal of Quality Technology}, {\bf 20}, 24--35.

\bibitem{nelson90a}
Nelson, W. (1990a), {\em Accelerated Testing: Statistical Models, Test
Plans, and Data Analyses}, New York: John
Wiley \& Sons.

\bibitem{nelson90b}
Nelson, W. (1990b), Hazard plotting of left truncated life data,
{\em Journal of Quality Technology}, {\bf 22},
230--238.

\bibitem{nelson95a} 
Nelson, W. (1995a), Confidence limits for recurrence data---applied to
cost or number of product repairs, {\em Technometrics}, {\bf 37},
147--157.

\bibitem{nelson95b} 
Nelson, W. (1995b), Defect initiation and growth---a general statistical
model \& data analysis, paper presented at the 2nd annual Spring
Research Conference, sponsored by the Institute of Mathematical
Statistics and the Physical and Engineering Section of the American
Statistical Association, Waterloo, Ontario, Canada, June 1995.

\bibitem{nelson95c}
Nelson, W. (1995c), Weibull prediction of a future number of failures,
paper presented at the 1995 Joint Statistical Meetings, Orlando, FL,
August 1995.

\bibitem{nelson98}
Nelson, W. (1998), Bibliography on accelerated test plans.
Available from the author, 739 Huntingdon Dr., Schenectady, NY 12309-2917.

\bibitem{nd89}
Nelson, W., and Doganaksoy, N.(1989), A computer program for an
estimate and confidence limits for the mean cumulative function for
cost or number of repairs of repairable products, TIS report 89CRD239,
General Electric Company Research and Development, Schenectady, NY.

\bibitem{nd95}
Nelson, W., and Doganaksoy, N. (1995), Statistical analysis of life or
strength data from specimens of various size using the
power-(log)normal model, Chapter 21 in {\em Recent Advances in
Life-Testing and Reliability}, N. Balakrishnan, Editor. Boca Raton,
FL: CRC Press.

\bibitem{nk76}
Nelson, W., and Kielpinski, T. (1976), Theory for optimum censored
accelerated life tests for normal and lognormal distributions, {\em
Technometrics}, {\bf 18}, 105--114.

\bibitem{nm78}
Nelson, W., and Meeker, W. (1978), Theory for optimum accelerated
censored life tests for Weibull and extreme value distributions, {\em
Technometrics}, {\bf 20}, 171--177.

\bibitem{ns81}
Nelson, W., and Schmee, J. (1981), Prediction limits for the last
failure of a (log) normal sample from early failures, {\em IEEE
Transactions  on Reliability, } {\bf R-30}, 461--463.


\bibitem{nt71}
Nelson, W., and Thompson, V. C. (1971), Weibull probability plots, {\em
Journal of Quality Technology}, {\bf 3}, 45--50.

\bibitem{neufelder}
Neufelder, A. (1993), {\em Ensuring Software Reliability}, New York:
Marcel Dekker.

\bibitem{nknw96}
Neter, J., Kutner, M. H., Nachtsheim, C. J., and Wasserman, W. (1996),
{\em Applied Linear Statistical Models} (Fourth Edition), Homewood,
IL: Richard D.  Irwin.

\bibitem{oconnor85}
O'Connor, P. D. T. (1985), {\em Practical Reliability Engineering}
(Second Edition), New York: John Wiley \& Sons.

\bibitem{oo80}
Odeh, R. E., and Owen, D. B. (1980), {\em Tables for Normal Tolerance
Limits, Sampling Plans, and Screening}, New York: Marcel Dekker.

\bibitem{om88}
Ostrouchov, G., and Meeker, W. Q. (1988), Accuracy of approximate
confidence bounds computed from interval censored Weibull and
lognormal data, {\em Journal of Statistical Computation and
Simulation}, {\bf 29}, 43--76.

\bibitem{owen90}
Owen, A. (1990), Empirical likelihood confidence regions, {\em The
Annals of Statistics}, {\bf 18}, 90--120.

\bibitem{parida91}
Parida, N. (1991), Reliability and life estimation from component
fatigue failures below the go~--~no~--~go fatigue limit, {\em Journal of
Testing and Evaluation}, {\bf 19}, 450--453.

\bibitem{pm97}
Pascual, F. G., and Meeker, W. Q. (1997), Regression analysis of
fatigue data with runouts based on a model with nonconstant
standard deviation and a fatigue limit parameter,
{\em Journal of Testing and Evaluation}, {\bf 25}, 292--301.

\bibitem{pm98a}
Pascual, F. G., and Meeker, W. Q. (1998a), Estimating fatigue curves
with the random fatigue--limit model.  Preprint, Department of
Statistics, Iowa State University, Ames, IA.

\bibitem{pm98b}
Pascual, F. G., and Meeker, W. Q. (1998b), The modified sudden death
test: planning life tests with a limited number of test
positions. Preprint, Department of Statistics, Iowa State
University, Ames, IA.

\bibitem{patel89}
Patel, J. K. (1989), Prediction intervals---a review, {\em
Communications in Statistics---Theory and Methods, } {\bf 18},
2393--2465.

\bibitem{pecht95}
Pecht, M., Editor (1995), {\em Product Reliability, Maintainability,
and Supportability Handbook}, Boca Raton, FL: CRC Press.

\bibitem{peck86}
Peck, D. S. (1986), Comprehensive model for humidity testing
correlation, {\em Proceedings of the IEEE International
Reliability Physics Symposium}, {\bf 24}, 44--50.

\bibitem{pz74}
Peck, D. S., and Zierdt, C. H. Jr.  (1974), The reliability of
semiconductor devices in the Bell System, {\em Proceedings of the
IEEE}, {\bf 62}, 185--211.

\bibitem{peto73}
Peto, R. (1973), Experimental survival curves for interval-censored
data, {\em Applied Statistics}, {\bf 22}, 86--91.

\bibitem{phadke89}
Phadke, M.~S. (1989), {\em Quality Engineering Using Robust Design},
Englewood Cliffs, NJ: Prentice Hall.

\bibitem{pb95a}
Pinheiro, J. C., and Bates, D. M. (1995a), Approximations to the
loglikelihood function in the nonlinear mixed effects model, 
{\em Journal of Computational and Graphical Statistics},  {\bf 4}, 12--35.

\bibitem{pb95b}
Pinheiro, J. C., and Bates, D. M. (1995b), Mixed effects models,
methods, and classes for S and SPLUS.  Department of Statistics,
University of Wisconsin.  Available from Statlib.

\bibitem{pollino89}
Pollino, E. (1989), {\em Microelectronic Reliability, Volume II,
Reliability, Integrity Assessment and Assurance}, Norwood, MA: Artech
House.

\bibitem{prentice74}
Prentice, R. L. (1974), A log gamma model and its maximum likelihood
estimation, {\em Biometrika}, {\bf 61}, 539--544.

\bibitem{prentice75}
Prentice, R. L. (1975), Discrimination among some parametric models,
{\em Biometrika}, {\bf 62}, 607--614.

\bibitem{proschan63}
Proschan, F. (1963), Theoretical explanation of observed decreasing
failure rate, {\em Technometrics}, {\bf 5}, 375--383.

\bibitem{rao73}
Rao, C. R. (1973), {\em Linear Statistical Inference and Its
Applications}, New York: John Wiley \& Sons.

\bibitem{ripley87}
Ripley, B. D. (1987), {\em Stochastic Simulation}, New York: John
Wiley \& Sons.

\bibitem{robinson83}
Robinson, J. A. (1983), Bootstrap confidence intervals in
location-scale models with progressive censoring, {\em Technometrics},
{\bf 25}, 179--187.

\bibitem{robinson95}
Robinson, J. A. (1995), Standard errors for the mean number of repairs
on systems from a finite population, Chapter 11 in {\em Recent
Advances in Life-Testing and Reliability}, N. Balakrishnan, Editor.
Boca Raton, FL: CRC Press.

\bibitem{rm91}
Robinson, J. A., and McDonald, G. C. (1991), Issues related to field
reliability and warranty data, in {\em Data Quality Control: Theory
and Pragmatics}, G. E. Liepins and V. R. R. Uppuluri, Editors. New York:
Marcel Dekker.

\bibitem{ross90}
Ross, G. J. S. (1990), {\em Nonlinear Estimation}, New York:
Springer-Verlag.
 
\bibitem{sas95}
SAS Institute Inc. (1995), {\em JMP User's Guide}, Version 3.1, 
Cary, NC: SAS Institute.

\bibitem{sas97}
SAS Institute Inc. (1997), SAS/QC Software: Changes and Enhancements 
for Release 6.12, Cary, NC: SAS Institute.

\bibitem{schinner96} 
Schinner, C. (1996), Accelerated Testing, Chapter 12 in
{\em Handbook of Reliability Engineering and Management} (Second
Edition), W. G. Ireson, C. F. Coombs, and R. Y.  Moss, Editors.  New
York: McGraw Hill.

\bibitem{sh79}
Schmee, J., and Hahn, G. J. (1979), A simple method for regression
analysis with censored data, {\em Technometrics}, {\bf 21}, 417--432.

\bibitem{sch86}
Schneider, H. (1986),
{\em Truncated and Censored Samples for Normal Populations},
New York: Marcel Dekker.

\bibitem{schrodinger15}
Schr\"odinger, E. (1915), Zur theorie der fall-und steigversuche an
teilchen mit Brownscher bewegung, {\em Physikalische Zeitschrift},
{\bf 16}, 289--295.

\bibitem{seber77}
Seber, G. A. F. (1977), {\em Linear Regression Analysis}, New York:
John Wiley \& Sons.

\bibitem{sw89}
Seber, G. A. F., and Wild, C. J. (1989), {\em Nonlinear Regression},
New York: John Wiley \& Sons.

\bibitem{severini91}
Severini, T. A. (1991), On the relationship between Bayesian and
Non-Bayesian interval estimates, {\em Journal of the Royal Statistical
Society B}, {\bf 53}, 611--618.

\bibitem{st95}
Shao, J., and Tu, D. (1995), {\em The Jackknife and Bootstrap}, New
York: Springer-Verlag.

\bibitem{sy79}
Shiomi, H., and Yanagisawa, T. (1979), On distribution parameter
during accelerated life test for a carbon film resistor, {\em
Bulletin of the Electrotechnical Laboratory}, {\bf 43}, 330--345
(in Japanese).

\bibitem{shooman}
Shooman, M. L. (1983), {\em Software Engineering: Design, Reliability,
and Management}, New York: McGraw Hill.

\bibitem{sing88a}
Singpurwalla, N. D. (1988a), An interactive PC-based procedure for
reliability assessment incorporating expert opinion and survival data,
{\em Journal of the American Statistical Association}, {\bf 83}, 43--51.

\bibitem{sing88b}
Singpurwalla, N. D. (1988b), Foundational issues in reliability and
risk analysis, {\em SIAM Review}, {\bf 30}, 264--282.

\bibitem{scg75}
Singpurwalla, N. D., Castellino, V. C., and Goldschen, D. Y. (1975),
Inference from accelerated life tests using Eyring type
re-parameterizations,  {\em Naval Research Logistics Quarterly}, {\bf
22}, 289--296.

\bibitem{sg92}
Smith, A. F. M.,  and Gelfand, A. E. (1992), Bayesian statistics
without tears: a sampling-resampling perspective, {\em The American
Statistician}, {\bf 46}, 84--88.

\bibitem{smith96}
Smith, J. S. (1996), Physics of failure, Chapter 14 in {\em Handbook
of Reliability Engineering and Management} (Second Edition), W. G.
Ireson, C. F. Coombs, and R. Y.  Moss, Editors.  New York: McGraw Hill.

\bibitem{smith85}
Smith, R. L. (1985), Maximum likelihood estimation in a class of
nonregular cases, {\em Biometrika}, {\bf 72}, 67--90.

\bibitem{sn87}
Smith, R. L., and Naylor, J. C. (1987),
A Comparison of maximum likelihood and bayesian estimators for the
three-parameter Weibull distribution,
{\em Applied Statistics}, {\bf 36}, 358--369.

\bibitem{snyder75}
Snyder, D. L. (1975), {\em Random Point Processes}, New York: John
Wiley \& Sons.

\bibitem{statsci1996}
S-plus, Statistical Sciences, Inc. (1996), {\em \splus User's Manual,
Volumes 1-2}, Version 3.4, Seattle: Statistical Sciences.

\bibitem{sprott73}
Sprott, D. A. (1973), Normal likelihoods and relation to a large
sample theory of estimation, {\em Biometrika}, {\bf 60}, 457--465.

\bibitem{starke94}
Starke, E. A., et al. (1996), {\em Accelerated Aging of Materials and
Structures}, Publication NMAB-479, Washington DC: National Academy Press.

\bibitem{so94}
Stuart A., and Ord, K. J. (1994), {\em Kendall's Advanced Theory of
Statistics}, New York: Edward Arnold.

\bibitem{sundararajan}
Sundararajan, C. (1991), {\em Guide to Reliability Engineering}, New
York: Van Nostrand Reinhold.

\bibitem{suzuki85}
Suzuki, K. (1985), Estimation method of lifetime based on the
record of failures during the warranty period, {\em Journal of the
American Statistical Association}, {\bf 80}, 66--72.

\bibitem{smy93}
Suzuki, K., Maki, K., and Yokogawa, S. (1993), An analysis of
degradation data of a carbon film and properties of the estimators, in
{\em Statistical Sciences and Data Analysis}, 501--511. K.~Matusita, M.~Puri,
and T.~Hayakawa, Editors.  Utrecht, Netherlands: VSP.

\bibitem{thatcher64}
Thatcher, A. R. (1964), Relationships between Bayesian and confidence
limits for prediction (with discussion), {\em Journal of the Royal
Statistical Society B}, {\bf 26}, 176--210.

\bibitem{thomas64}
Thomas, R. (1964), When is a life test truly accelerated? {\em
Electronic Design}, January 6,  64--71.

\bibitem{tg75}
Thomas, D. R., and Grunkemeier, G. L. (1975), Confidence interval
estimation of survival probabilities for censored data, {\em Journal
of the American Statistical Association}, {\bf 70}, 865--871.

\bibitem{thompson81} 
Thompson, W. A. (1981), On the foundations of reliability, {\em
Technometrics}, {\bf 23}, 1--14.

\bibitem{thompson88} 
Thompson, W. A. (1988), {\em Point Process Models with Applications to
Safety and Reliability}, London: Chapman \& Hall.

\bibitem{tsm85}
Titterington, D. M., Smith, A. F. M., and Makov, U. E. (1985), {\em
Statistical Analysis of Finite Mixture Distributions}, New York: John
Wiley \& Sons.

\bibitem{tt95}
Tobias, P. A., and Trindade, D. C. (1995), {\em Applied Reliability}
(Second Edition), New York: Van Nostrand Reinhold.

\bibitem{tomsky82}
Tomsky, J. (1982), Regression models for detecting reliability
degradation, {\em 1982 Proceedings of the Annual Reliability and
Maintainability Conference},  238--244, New York: Institute of
Electrical and Electronics Engineers.

\bibitem{trindade91}
Trindade, D. C. (1991), Can burn-in screen wearout mechanisms?
Reliability models of defective subpopulations---a case study, {\em
Proceedings of the IEEE International Reliability Physics
Symposium}, {\bf 29}, 260--263.

\bibitem{thc95}
Tseng, S. T., Hamada, M., and Chiao, C. H. (1995), Using degradation
data from a factorial experiment to improve fluorescent lamp
reliability, {\em Journal of Quality Technology}, {\bf 27}, 363--369.

\bibitem{tw97}
Tseng, S. T., and Wen, Z. C. (1997), Step-stress accelerated
degradation analysis for highly reliable products, Institute of
Statistics, National Tsing-Hua University.

\bibitem{ty97}
Tseng, S. T., and Yu, H. F. (1997), A termination rule for
degradation experiment, {\em IEEE Transactions on Reliability}, {\bf
46}, 130--133.

\bibitem{turn76}
Turnbull, B. W. (1976), The empirical distribution function with
arbitrary grouped, censored, and truncated data, {\em Journal of the
Royal Statistical Society}, {\bf 38}, 290--295.

\bibitem{tustin90}
Tustin, W. (1990), Shake and bake the bugs out, {\em Quality
Progress}, September, 61--64.

\bibitem{tweedie45}
Tweedie, M. C. K. (1945), Inverse statistical variates, {\em Nature},
{\bf 155}, 453.

\bibitem{ueda96}
Ueda, O. (1996), {\em Reliability and Degradation of III-V Optical Devices}, 
Norwood, MA: Artech House.

\bibitem{ushakov94}
Ushakov, I., Editor (1994), {\em Handbook of Reliability
Engineering}, New York: John Wiley \& Sons.

\bibitem{vm90}
Vander Wiel, S. A., and Meeker, W. Q. (1990), Accuracy of approximate
confidence bounds using censored Weibull regression data from
accelerated life tests, {\em IEEE Transactions on Reliability}, {\bf
R-39}, 346--351.

\bibitem{viertl88}
Viertl, R. (1988), {\em Statistical Methods for Accelerated Life
Testing}, G\"{o}ttingen: Vandenhoeck and Ruprecht.

\bibitem{wald47}
Wald, A. (1947), {\em Sequential Analysis, } New York: John Wiley \&
Sons.
  
\bibitem{wang91}
Wang, C. J. (1991), Sample size determination of bogey tests without
failures, {\em Quality and Reliability Engineering International},
{\bf 7}, 35--38.
 
\bibitem{wcsc86}
Weis, E. A., Caldararu, D., Snyder, M. M., and Croitoru, N. (1986),
Investigating reliability attributes of silicon photodetectors, {\em
Microelectronics and Reliability}, {\bf 26}, 1099--1110.

\bibitem{ws72}
Weston, R., and Schwarz, H. A. (1972), {\em Chemical Kinetics},
Englewood Cliffs, NJ: Prentice-Hall.

\bibitem{wm90}
Weston, S. A., and Meeker, W. Q. (1990), Coverage probabilities of
nonparametric simultaneous confidence bands for a survival function,
{\em Journal of Statistical Simulation and Computation}, {\bf 32},
83--97.
 
\bibitem{wgh62a}
Wilk, M. B., Gnanadesikan, R., and Huyett, M. J. (1962a), Probability
plots for the gamma distribution, {\em Technometrics}, {\bf 4}, 1--20.

\bibitem{wgh62b}
Wilk, M. B., Gnanadesikan, R., and Huyett, M. J. (1962b), Estimation
of the parameters of the gamma distribution
using order statistics, {\em Biometrika}, {\bf 49}, 525--545.

\bibitem{wood72}
Woodroofe, M. (1972), Maximum likelihood estimate of a translation
parameter of a truncated distribution, {\em Annals of Mathematical
Statistics}, {\bf 43}, 113--122.

\bibitem{wood74}
Woodroofe, M. (1974), Maximum likelihood estimate of a translation
parameter of a truncated distribution II, {\em Annals of Statistics},
{\bf 2}, 474--488.

\bibitem{yanagisawa1997}
Yanagisawa, T. (1997), Estimation of the degradation of amorphous
silicon solar cells, {\em Microelectronics and Reliability}, {\bf 37},
549--554.

\bibitem{yokobori51}
Yokobori, T. (1951), Fatigue fracture in steel, {\em Journal of the
Physical Society of Japan}, {\bf 6}, 81--86.

\bibitem{yt98}
Yu, H. F., and Tseng, S. T. (1998), On-line procedure for
terminating an accelerated degradation test. To appear in {\em
Statistica Sinica}.

\bibitem{zelen59}
Zelen, M. (1959), Factorial experiments in life testing, {\em
Technometrics}, {\bf 1}, 269--288.

\end{thebibliography}

