\chapter*{Acknowledgments}

A number of individuals provided helpful comments on all or part of
draft versions of this book.  In particular, we would like to
acknowledge Chuck Annis, Necip Doganaksoy, Shuen-Lin Jeng, Gerry
Hahn, Joseph Lu, Michael LuValle, Enid Martinets, Peter Morse,
Ernest Scheuer, Ananda Sen, David Steinberg, Mark Vandeven, and a
number of anonymous reviewers.  Over the years, we have benefited
from numerous technical discussions with Vijay Nair. Vijay provided
a number of suggestions that substantially improved several parts of
this book. We would also like to acknowledge our students (too
numerous to mention by name) for their penetrating questions, high
level of interest, and for their useful suggestions for improving
our courses.

We would like to make special acknowledgment to Wayne Nelson who
gave us detailed feedback on earlier versions of most of the
chapters of this book.  Additionally, much of our knowledge in this
area has it roots in our interactions with Wayne and Wayne's
outstanding books and other publications in the area of reliability
and reliability data analysis.

Parts of this book were written while the first author was visiting
the Department of Statistics and Actuarial Science at the University
of Waterloo and the Department of Experimental Statistics at
Louisiana State University. Support and use of facilities during these
visits is gratefully acknowledged.

We greatly benefited from facilities, traveling support, and
encouragement from Jeff Hooper and Mich\`{e}le Boulanger at AT\&T Bell
Laboratories, Gerald Hahn at General Electric Corporate Research and
Development Center, Enrique Villa and Victor P\'{e}rez Abreu at CIMAT
in Guanajuato, M\'{e}xico, Guido E.\ Del Pino at Universidad
Cat\'{o}lica in Santiago, Chile, H\'{e}ctor Allende at Universidad
T\'{e}cnica Federico Santa Mar\'{\i}a in Valparaiso, Chile, Yves L.
Grize in Basel, Switzerland, and Steve Zayac, Ford Motor Company.

Dean L. Isaacson, Head, Department of Statistics, Iowa State
University, Lynn  R.\ LaMotte, former Head, and E.\ Barry Moser,
Interim Head, Department of Experimental Statistics, Louisiana State
University, provided helpful encouragement and support to both
authors. We would also like to thank our secretaries Denise Riker
and Elaine Miller for excellent support and assistance while writing
this book.

Finally, we would like to thank our wives and children for their
love, patience, and understanding during the recent years in which
we worked most weekends and many evenings to complete this project.





