---
output: html_document
---

__It is possible for a continuous cdf to be constant over some intervals of time__

a. __Give an example of a physical situation that would result in a cdf, $F(t)$ that is constant over some values of $t$.__

`r teachingSMRD:::solution()$begin`
Suppose that a product's lifetime is measured using clock time, but the units do not operate during some portion of each.  During the time that the product is not operating the probability of failure, as expressed by the pdf, would be zero.  As result, there would be no accumulated probability of failure during the off-times and the cdf would remain constant (i.e. flat). 

As an example, suppose we are modelling the reliability of a light bulb using days to failure as our time-scale.  Now suppose that the bulb is switched off for 12 hours each day.  Since the bulb can only fail during the operating periods the cdf for the failure time distribution will be flat for half of each day.  The occurence of flat spots in the cdf of a failure time distribution suggest that time-scale used to measure life is inappropriate. A more appropriate time-scale to use for modelling time to failure of the light bulbs in our example, might be true operating hours instead of calendar days.  Of course, choosing to use true operating hours as a measure of how life is consumed requires that we have the ability to monitor the total number of hours for which the light bulbs have been operating.
`r teachingSMRD:::solution()$end`

b. __Sketch such a cdf and its corresponding pdf__

`r teachingSMRD:::solution()$begin`
Suppose that the lifetime of the lightbulbs discussed in part (a) follows a $Lognor(5,.85)$ distribution (this is a poor assumption since a lightbulb's liftime is one of the few things that is known to follow an exponential distribution).  Under this assumption, if the lightbulbs were used continuously (24-hrs/day) the respective pdf and cdf plots would appear as shown in the figure below.

```{r,  fig.cap="Pdf (left) and cdf curves for a lognormal(5,.85) distribution assuming continuous (24-hr) usage", fig.width=9, fig.height=5, echo=2:9}
## Configure plot area parameters
par(lwd = 2, las = 1, mgp = c(2.5,.25,0), tcl = .3)

## Set 1-row X 2-col plot layout
layout(mat = matrix(1:2, 1,2))

curve(dlnorm(x,5,.85),
      from = 0,
      to = 480,
      xlab = "Hours", 
      ylab = "f(t)")
curve(plnorm(x,5,.85),
      from = 0,
      to = 480,
      xlab = "Hours", 
      ylab = "F(t)")
```

If the light bulbs did not operate continuously, but were instead only used in 12-hour increments the cdf and pdf plots would appear as shown in following figure.  (Note, these figure are based on the assumption that clock time -- not operating time -- is an appropriate time scale to model time to failure. 

```{r,  fig.cap="Pdf (left) and cdf curves for a lognormal(5,.85) distribution assuming half-day usage (12-hr intervals)" ,echo=1:15, fig.width=9, fig.height=5}
points <- seq(0,480,12)
p <- points[c(1,rep(2:70,each = 2),71)]
Z <- function(a) rep(plnorm(p[a],5,.85),13)
z <- rep(0,13)

# Set up plot area
par(lwd = 2, las = 1, mgp = c(2.5,.25,0), tcl = .3)

## Set 1-row X 2-col plot layout
layout(mat = matrix(1:2, 1,2))

plot(
c(p[ 1]:p[ 2],p[ 3]:p[ 4],p[ 5]:p[ 6],p[ 7]:p[ 8],p[ 9]:p[10],
  p[11]:p[12],p[13]:p[14],p[15]:p[16],p[17]:p[18],p[19]:p[20],
  p[21]:p[22],p[23]:p[24],p[25]:p[26],p[27]:p[28],p[29]:p[30],
  p[31]:p[32],p[33]:p[34],p[35]:p[36],p[37]:p[38],p[39]:p[40],
  p[41]:p[42],p[43]:p[44],p[45]:p[46],p[47]:p[48],p[49]:p[50],
  p[51]:p[52],p[53]:p[54],p[55]:p[56],p[57]:p[58],p[59]:p[60],
  p[61]:p[62],p[63]:p[64],p[65]:p[66],p[67]:p[68],p[69]:p[70]),
    
c(dlnorm(p[ 1]:p[ 2],5,.85),z,dlnorm(p[ 3]:p[ 4],5,.85),z,
  dlnorm(p[ 5]:p[ 6],5,.85),z,dlnorm(p[ 7]:p[ 8],5,.85),z,
  dlnorm(p[ 9]:p[10],5,.85),z,dlnorm(p[11]:p[12],5,.85),z, 
  dlnorm(p[13]:p[14],5,.85),z,dlnorm(p[15]:p[16],5,.85),z,
  dlnorm(p[17]:p[18],5,.85),z,dlnorm(p[19]:p[20],5,.85),z,
  dlnorm(p[21]:p[22],5,.85),z,dlnorm(p[23]:p[24],5,.85),z,
  dlnorm(p[25]:p[26],5,.85),z,dlnorm(p[27]:p[28],5,.85),z,
  dlnorm(p[29]:p[30],5,.85),z,dlnorm(p[31]:p[32],5,.85),z,
  dlnorm(p[33]:p[34],5,.85),z,dlnorm(p[35]:p[36],5,.85)),
type="l",
xlab = "Hours", 
ylab = "f(t)")

plot(
c(p[ 1]:p[ 2],p[3] :p[ 4],p[ 5]:p[ 6],p[ 7]:p[ 8],p[ 9]:p[10],
  p[11]:p[12],p[13]:p[14],p[15]:p[16],p[17]:p[18],p[19]:p[20],
  p[21]:p[22],p[23]:p[24],p[25]:p[26],p[27]:p[28],p[29]:p[30],
  p[31]:p[32],p[33]:p[34],p[35]:p[36],p[37]:p[38],p[39]:p[40],
  p[41]:p[42],p[43]:p[44],p[45]:p[46],p[47]:p[48],p[49]:p[50],
  p[51]:p[52],p[53]:p[54],p[55]:p[56],p[57]:p[58],p[59]:p[60],
  p[61]:p[62],p[63]:p[64],p[65]:p[66],p[67]:p[68],p[69]:p[70]),
    
c(plnorm(p[ 1]:p[ 2],5,.85),Z( 2),plnorm(p[ 3]:p[ 4],5,.85),Z( 4),
  plnorm(p[ 5]:p[ 6],5,.85),Z( 6),plnorm(p[ 7]:p[ 8],5,.85),Z( 8),
  plnorm(p[ 9]:p[10],5,.85),Z(10),plnorm(p[11]:p[12],5,.85),Z(12), 
  plnorm(p[13]:p[14],5,.85),Z(14),plnorm(p[15]:p[16],5,.85),Z(16),
  plnorm(p[17]:p[18],5,.85),Z(18),plnorm(p[19]:p[20],5,.85),Z(20),
  plnorm(p[21]:p[22],5,.85),Z(22),plnorm(p[23]:p[24],5,.85),Z(24),
  plnorm(p[25]:p[26],5,.85),Z(26),plnorm(p[27]:p[28],5,.85),Z(28),
  plnorm(p[29]:p[30],5,.85),Z(30),plnorm(p[31]:p[32],5,.85),Z(32),
  plnorm(p[33]:p[34],5,.85),Z(34),plnorm(p[35]:p[36],5,.85)),
type = "l", 
xlab = "Hours",
ylab = "F(t)")
```
`r teachingSMRD:::solution()$end`

d. __Think about and explain the relationship between the occurrence of flat spots in a cdf and the choice of an approriate time scale on which to define life.__

`r teachingSMRD:::solution()$begin`
Continuing with the lightbulb example, the flat spots in the cdf shown in figure above occur because the time scale used to model time to failure did not accurately represent the way that life was consumed.  Had "operating hours" been used instead of "clock-hours", the cdf plot would appear the same as the first plot shown above.
`r teachingSMRD:::solution()$end`
