---
output: html_document
---

__The coefficient of variation, $\gamma_{2},$ is a useful scale-free measure of relative variability for a random variable.__

a. __Derive an expression for the coefficient of variation for the Weibull distribution__

`r teachingSMRD:::solution()$begin`
Using the results from Exercise 4.5, we can express the Weibull coefficient of variation as

$$
\gamma_2=\sqrt{\frac{\sigma^2}{\mu^2}}=\left[\Gamma\left(1+\frac{2}{\beta}\right)\Gamma^{-2}\left(1+\frac{1}{\beta}\right)-1\right]^{0.5}
$$
`r teachingSMRD:::solution()$end`

b. __Compute $\gamma_{2}$ for all combinations of $\beta=0.5,1,3,5$ and $\eta=50,100.$  Also, draw (or use the computer to draw) a graph of the Weibull pdfs for the same combinations of parameters.__

`r teachingSMRD:::solution()$begin`
As seen in (a), the value of $\gamma_2$ for the Weibull distribution depends only on the value of the shape parameter $\beta$.  For a given value of $\beta$ the value of $\gamma_2$ can be found using the following R code

```{r,echo=TRUE}
beta <- c(0.5,1,3,5)

data.frame(CoV = (gamma(1+2/beta)*(gamma(1+1/beta))^-2-1)^0.5,
           row.names =c("eta(0.5) =",
                        "eta(1.0) =",
                        "eta(3.0) =",
                        "eta(5.0) =") )
```

The plots of the respective pdfs may then be found using

```{r,echo=TRUE,fig.align='center', fig.cap="Plots of various Weibull densities"}
beta <- rep(c(0.5,1,3,5),2)
eta <- rep(c(50,100),each = 4)

par(family="serif")

for(i in 1:8) {
curve(dweibull(x, shape = beta[i], scale = eta[i]), 
      add = ifelse(i==1,F,T), 
      col = i, 
      lty = i, 
      lwd = 2.5, 
      xlim = c(0,100), ylim = c(0,0.04),
      xlab = 't', ylab = 'f(t)')
}
legend("topright", 
       legend = c(expression(beta*','*eta=="0.5,50"),
                  expression(beta*','*eta=="1.0,50"),
                  expression(beta*','*eta=="3.0,50"),
                  expression(beta*','*eta=="5.0,50"),
                  expression(beta*','*eta=="0.5,100"),
                  expression(beta*','*eta=="1.0,100"),
                  expression(beta*','*eta=="3.0,100"),
                  expression(beta*','*eta=="5.0,100")),
         lty = 1:8,
         lwd = 2,
         col = 1:8,
         cex = 0.75,
         bty = "n",
         xjust = 0.5)
```
`r teachingSMRD:::solution()$end`

c. __Explain the effect that changes in $\eta$ and $\beta$ have on the shape of the Weibull density and the effect that they have on $\gamma_{2}.$__

`r teachingSMRD:::solution()$begin`
In problem 4.7 we discussed the effect changing the shape parameter $\beta$ has on the shape of the Weibull distribution, thus we need not discuss it again here.  Increasing the scale parameter $\eta$ or "characteristic life" increases the skewness of the distriution by shifting the $0.632$ quantile to the right.

The coefficient of variation $\gamma_{2}$ is a function of the shape parameter alone.  The figure below illustrates the effect of changing $\beta$

```{r,echo=FALSE,fig.align='center', fig.cap="Weibull coefficient of variation vs shape parameter"}
CoV <- function(beta) { ((((gamma(1+2/beta))/(gamma(1+1/beta))^2)-1)^0.5)}

par(family="serif", font = 2)

curve(CoV(x), 
      xlim = c(0.5,25),
      ylab = expression(gamma[2]),
      xlab = expression(beta),
      lwd = 2,
      las = 1)
```
`r teachingSMRD:::solution()$end`
