---
title: "`r SMRD:::info('book')`"
subtitle: "`r SMRD:::info('chapter13')`"
author: "`r SMRD:::info('authors')`"
date: "`r format(Sys.Date(), '%d %B %Y')`"
footer: "`r paste('SMRD: ', SMRD:::info('chapter13'))`"
output:
  slidy_presentation:
    smart: false
    fig_caption: yes
runtime: shiny
graphics: yes
---

```{r echo=FALSE,message=FALSE, warning=FALSE}
assign('dynamic', TRUE, envir = sys.frame(),immediate = TRUE)
source('scripts/R/setup.R')
```

# CHAPTER OVERVIEW

```{r}
shiny::includeCSS('scripts/css/flat-slidy.css')
shiny::includeScript(path = "scripts/js/audiojs/audiojs/audio.min.js",
                     "audiojs.events.ready(function() {audiojs.createAll();});")
shiny::includeScript("scripts/js/jkf-scroll.js")
```

## This chapter explains

- Some useful degradation models

- The connection between degradation models and failure-time models

- How degradation measures, when available, can be used to advantage in estimating reliability

- Methods for data analysis and reliability inference with degradation data

- The differences between degradation data analysis and traditional failure-time data analysis

- A simple method for degradation analysis that can be used in certain situations

# 13.1 - INTRODUCTION

# 13.2 - MODELS FOR DEGRADATION DATA

#  13.2.1 - Degradation data

#  13.2.2 - Degradation leading to failure

#  13.2.3 - Models for variation in degradation and failure times

#  13.2.4 - Limitations of degradation data

#  13.2.5 - General degradation path model

#  13.2.6 - Degradation model parameters

# 13.3 - ESTIMATION OF DEGRADATION MODEL PAMARETERS

# 13.4 - MODELS RELATING DEGRADATION AND FAILURE

#  13.4.1 - Soft failures: specified degradation level


#  13.4.2 - Hard failures: joint distribution of degradation and failure level

# 13.5 - EVALUATION OF $F(t)$

#  13.5.1 - Analytical solution for $F(t)$

#  13.5.2 - Numerical evaluation of $F(t)$

#  13.5.3 - Monte Carlo evaluation of $F(t)$

# 13.6 - ESTIMATION OF $F(t)$

# 13.7 - BOOTSTRAP CONFIDENCE INTERVALS

# 13.8 - COMPARISON WITH TRADITIONAL FAILURE-TIME ANALYSIS

# 13.9 - APPROXIMATE DEGRADATION ANALYSIS

#  13.9.1 - Simple linear path


#  13.9.2 - Simple linear path through the origin


#  13.9.3 - Comments on the approximate degradation analysis

<link rel='stylesheet' type='text/css' href='css/flat-slidy.css'><script src='js/audiojs/audiojs/audio.min.js'></script><script> audiojs.events.ready(function() {audiojs.createAll();}); </script><script src='js/jkf-scroll.js'></script>