---
title: "`r SMRD:::info('book')`"
subtitle: "`r SMRD:::info('chapter18')`"
author: "`r SMRD:::info('authors')`"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  slidy_presentation:
    smart: false
    fig_caption: yes
footer: "`r paste('SMRD: ', SMRD:::info('chapter18'))`"
runtime: shiny
graphics: yes
---

```{r intro, echo=FALSE,message=FALSE, warning=FALSE}
assign('dynamic', TRUE, envir = sys.frame(),immediate = TRUE)
source('scripts/R/setup.R')
```

# CHAPTER OVERVIEW

```{r}
shiny::includeCSS('scripts/css/flat-slidy.css')
shiny::includeScript(path = "scripts/js/audiojs/audiojs/audio.min.js",
                     "audiojs.events.ready(function() {audiojs.createAll();});")
shiny::includeScript("scripts/js/jkf-scroll.js")
```

## This chapter explains

- Motivation and applications of accelerated testing

- Connections between degradation and physical failure

- Models for temperature acceleration

- Models for voltage and pressure acceleration

- How to compute time-acceleration factors

- Other accelerated test models and their assumptions 

# 18.1 - INTRODUCTION

# 18.1.1 - Motivation 


# 18.1.2 - Different types of acceleration


# 18.1.3 - Types of responses 


# 18.1.4 - Methods of acceleration 


# 18.1.5 -Acceleration models 


# 18.2 - USE-RATE ACCELERATION

# 18.3 - TEMPERATURE ACCELERATION

# 18.3.1 - Arrhenius time-acceleration factor


# 18.3.2 - Eyring time-acceleration factor


# 18.3.3 - Reactio-rate acceleration for a nonlinear degradation path model 


# 18.3.4 - Acceleration for a linear degradation path model 


# 18.3.5 - Acceleration for parallel chemical reations 

#  18.4 - VOLTAGE AND VOLTAGE-STRESS ACCELERATION

# 18.4.1 - Voltage acceleration mechanisms


# 18.4.2 - The inverse power relationship


# 18.4.3 - Physical motivation for the inverse power relationship for voltage-stress acceleration


# 18.4.4 - Other inverse power relationships 

# 18.5 - ACCELERATION MODELS WITH MORE THAN ONE ACCELERATING VARIABLE

# 18.5.1 - Generalized Eyring relationship


# 18.5.2 - Temperature-voltage relationship


# 18.5.3 - Temperature-current density acceleration


# 18.5.4 - Temperature-humidity acceleration


# 18.6 - GUIDELINE FOR THE USE OF ACCELERATION MODELS

```{r}

```