---
title: "`r SMRD:::info('book')`"
subtitle: "`r SMRD:::info('chapter14')`"
author: "`r SMRD:::info('authors')`"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  slidy_presentation:
    smart: false
    fig_caption: yes
footer: "`r paste('SMRD: ', SMRD:::info('chapter14'))`"
runtime: shiny
graphics: yes
---

```{r intro, echo=FALSE,message=FALSE, warning=FALSE}
assign('dynamic', TRUE, envir = sys.frame(),immediate = TRUE)
source('scripts/R/setup.R')
```

# CHAPTER OVERVIEW

```{r}
shiny::includeCSS('scripts/css/flat-slidy.css')
shiny::includeScript(path = "scripts/js/audiojs/audiojs/audio.min.js",
                     "audiojs.events.ready(function() {audiojs.createAll();});")
shiny::includeScript("scripts/js/jkf-scroll.js")
```

## This chapter explains

- The use of Bayesian statistical methods to combine "prior" information with data to make inferences
- The relationship between Bayesian methods and the likelihood methds used in earlier chapters
- Sources of prior information
- Useful statistical and numerical methods for Bayesian analysis
- Bayesian methods for estimating reliability
- Bayesian methods for prediction
- The dangers of using "wishful thinking" or expectations as prior information

# 14.1 - INTRODUCTION

# 14.2 - USING BAYES'S RULE TO UPDATE PRIOR INFORMATION

#  14.2.1 Notation

#  14.2.2 Bayes's rule

# 14.3 - PRIOR INFORMATION AND DISTRIBUTIONS

#  14.3.1 Noninformative (diffuse) prior distributions

#  14.3.2 Using past data to specify a prior distribution

#  14.3.3 Expert opinion and eliciting prior information

# 14.4 - NUNERICAL METHODS FOR COMBINING PRIOR INFORMATION WITH A LIKELOHOOD

#  14.4.1 Numerical integration methods for computing the posterior distribution of $\theta$

#  14.4.2 Simulation-based methods for computing the posterior distribution of $\theta$

# 14.5 - USING THE POSTERIOR DISTRIBUTION FOR ESTIMATION

#  14.5.1 - Bayesian point estimation

#  14.5.2 - Bayesian interval estimation

# 14.6 - BAYESIAN PREDICTION

#  14.6.1 - Bayesian posterior predictive distribution

#  14.6.2 - Approximating posterior predictive distributions

#  14.6.3 - Prediction of an observation from a log-location-scale distribution

#  14.6.4 - Posterior predictive distribution for the $k$th failure from a future sample of size $m$

# 14.7 - PRACTICAL ISSUES IN THE APPLICATION OF BAYESIAN METHODS

#  14.7.1 - Comparison between Bayesian and likelihood/frequentist statistical methods

#  14.7.2 - Cautions on the use of prior information


```{r}

```