---
title: "`r SMRD:::info('book')`"
subtitle: "`r SMRD:::info('chapter12')`"
author: "`r SMRD:::info('authors')`"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  slidy_presentation:
    smart: false
    fig_caption: yes
footer: "`r paste('SMRD: ', SMRD:::info('chapter12'))`"
runtime: shiny
graphics: yes
---

```{r intro, echo=FALSE,message=FALSE, warning=FALSE}
assign('dynamic', TRUE, envir = sys.frame(),immediate = TRUE)
source('scripts/R/setup.R')
```

# CHAPTER OVERVIEW

```{r}
shiny::includeCSS('scripts/css/flat-slidy.css')
shiny::includeScript(path = "scripts/js/audiojs/audiojs/audio.min.js",
                     "audiojs.events.ready(function() {audiojs.createAll();});")
shiny::includeScript("scripts/js/jkf-scroll.js")
```

## This chapter explains

- Important reliability-related applications of prediction

- The difference between probability prediction and statistical prediction

- Methods for computing predictions and prediction bounds for future failure times

- Methods for computing predictions and prediction bounds for the number of failures in a future time interval

# 12.1 - INTRODUCTION

#  12.1.1 - Motivation and prediction problems
 

#  12.1.2 - Model
 

#  12.1.3 - Data
 

# 12.2 - PROBABILITY PREDICTION INTERVALS ($\theta$ GIVEN)


# 12.3 - STATISTICAL PREDICTION INTERVAL ($\theta$ ESTIMATED)

#  12.3.1 - Coverage probability concepts
 

#  12.3.2 - Relationship between 1-sided & 2-sided prediction intervals
 

#  12.3.3 - Naive method for computing a statistical prediction interval
 

# 12.4 - THE (APPROXIMATE) PIVOTAL METHOD FOR PREDICTION INTERVALS

#  12.4.1 - Type II (failure) censoring
 

#  12.4.2 - Type I (time) censoring
 

# 12.5 - PREDICTION IN SIMPLE CASES

#  12.5.1 - Complete samples from a lognormal distribution
 

#  12.5.2 - Complete of Type II censored samples from an exponential distribution
 

# 12.6 - CALIBRATING NAIVE STATISTICAL PREDICTION BOUNDS

#  12.6.1 - Calibration by simulation of the sampling/prediction process
 

#  12.6.2 - Calibration by averaging conditional coverage probabilities
 

# 12.7 - PREDICTION OF FUTURE FAILURES FROM A SINGLE GROUP OF UNITS IN THE FIELD


# 12.8 - PREDICTION OF FUTURE FAILURES FROM MULTIPLE GROUPS WITH STAGGERED ENTRY

```{r}

```