---
title: "`r SMRD:::info('book')`"
subtitle: "`r SMRD:::info('chapter13')`"
author: "`r SMRD:::info('authors')`"
date: "`r format(Sys.Date(), '%d %B %Y')`"
footer: "`r paste('SMRD: ', SMRD:::info('chapter13'))`"
output:
  slidy_presentation:
    smart: false
    fig_caption: yes
    df_print: paged
graphics: yes
---

```{r echo=FALSE,message=FALSE, warning=FALSE}
source(jkf::SCRIPT('setup.R'))
library(SMRD)
```

# CHAPTER OVERVIEW

```{r}
shiny::includeCSS(jkf::SCRIPT('flat-slidy.css'))
shiny::includeScript(jkf::SCRIPT("jquery.min.js"))
shiny::includeScript(jkf::SCRIPT("jkf-scroll.js"))
```

## Introduction

- Many modern products are highly reliable due better materials and manufacturing processes

- Testing high reliability systems produce very few failure observations (if any at all)

- Options for testing high-reliability systems

    1. Generate failures in less time by running the test in an elevated environment (accelerated life testing)
    2. Use prior information from similar systems or subject matter experts in conjuction with the test data (Bayesian reliability analysis)
    3. Use degradation measures as a physics-based proxy for time to failure (degradation data analysis)

## This chapter explains

- Some useful degradation models

- The connection between degradation models and failure-time models

- How degradation measures, when available, can be used to advantage in estimating reliability

- Methods for data analysis and reliability inference with degradation data

- The differences between degradation data analysis and traditional failure-time data analysis

- A simple method for degradation analysis that can be used in certain situations

# 13.2 - MODELS FOR DEGRADATION DATA

#  13.2.1 - Degradation data

#  13.2.2 - Degradation leading to failure

#  13.2.3 - Models for variation in degradation and failure times

#  13.2.4 - Limitations of degradation data

#  13.2.5 - General degradation path model

#  13.2.6 - Degradation model parameters

# 13.3 - ESTIMATION OF DEGRADATION MODEL PAMARETERS

# 13.4 - MODELS RELATING DEGRADATION AND FAILURE

#  13.4.1 - Soft failures: specified degradation level


#  13.4.2 - Hard failures: joint distribution of degradation and failure level

# 13.5 - EVALUATION OF $F(t)$

#  13.5.1 - Analytical solution for $F(t)$

#  13.5.2 - Numerical evaluation of $F(t)$

#  13.5.3 - Monte Carlo evaluation of $F(t)$

# 13.6 - ESTIMATION OF $F(t)$

# 13.7 - BOOTSTRAP CONFIDENCE INTERVALS

# 13.8 - COMPARISON WITH TRADITIONAL FAILURE-TIME ANALYSIS

# 13.9 - APPROXIMATE DEGRADATION ANALYSIS

#  13.9.1 - Simple linear path


#  13.9.2 - Simple linear path through the origin


#  13.9.3 - Comments on the approximate degradation analysis

<link rel='stylesheet' type='text/css' href='css/flat-slidy.css'><script src='js/audiojs/audiojs/audio.min.js'></script><script> audiojs.events.ready(function() {audiojs.createAll();}); </script><script src='js/jkf-scroll.js'></script>
