---
output: html_document
---

```{r}
library(jkf)
library(shiny)
```

A new, second-generation system (gen2) is being developed to deliver improved performance over the first-generation system (gen1).  The both the gen1 and gen2 systems contain the three-component sub-assembly shown below.

```{r, echo=FALSE}
shiny::img(src = system.file('images','simple-system.png', package = 'jkf'), 
           width = '60%', 
           align = 'middle')
```

Over the life of the gen1 system, the sub-assembly was highly reliable.  However, the gen2 system will operate under higher mechanical and electrical loads.  Management is therefore concerned that the sub-assembly is not rated appropriately and will be less reliable in the gen2 system.  Engineers have proposed prior distributions

```{r}
Number = c(1:2)
Component = c('$WEIB(\\eta, \\beta)$', '$EXP(\\theta)$')
df <- data.frame(Number,Component)
colnames(df) <- c('Component Number', 'Componenet Failure Distribution')
knitr::kable(df, format = 'pandoc')
```

The engineers who have extensive experience on the previous generation system


