---
title: "`r SMRD:::info('book')`"
subtitle: "`r SMRD:::info('chapter19')`"
author: "`r SMRD:::info('authors')`"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  slidy_presentation:
    smart: false
    fig_caption: yes
footer: "`r paste('SMRD: ', SMRD:::info('chapter19'))`"
runtime: shiny
graphics: yes
---

```{r intro, echo=FALSE,message=FALSE, warning=FALSE}
assign("dynamic",TRUE,envir = sys.frame(),immediate = TRUE)
assign("publish",TRUE,envir = sys.frame(),immediate = TRUE)

library(SMRD)
library(xtable)
library(jkf)

knitr::opts_chunk$set(message = FALSE, warning = FALSE, echo=FALSE)
```

# CHAPTER OVERVIEW

```{r}
shiny::includeCSS('scripts/css/flat-slidy.css')
shiny::includeScript(path = "scripts/js/audiojs/audiojs/audio.min.js",
                     "audiojs.events.ready(function() {audiojs.createAll();});")
shiny::includeScript("scripts/js/jkf-scroll.js")
```

## This chapter explains

- Nonparametric and graphical methods for presenting and analyzing accelerated life test (ALT) data

- Likelihood methods for analyzing right-censored data from an ALT with a single accelerating factor

- Analysis of other kinds of ALT experiments including experiments yielding interval data and experiments with two accelerating variables

- Some other common forms of accelerated testing

- Some potential pitfalls of accelerated testing 

# 19.1 - INTRODUCTION

# 19.1.1 - Accelerated life test models

<div id="play"><audio controls src="media/audio/test-slide.m4a"></audio></div>

# 19.1.2 - Strategy for analyzing ALT data


# 19.2 - ANALYZING OF SINGLE-VARIABLE ALT DATA

# 19.2.1 - Scatter plot of ALT data


# 19.2.2 - Multiple probability plot of noonparametric cdf estimates at individual levels of the accelerating variable


# 19.2.3 - Adding ML estimates at each level of the accelerating variables to a multiple probability plot


# 19.2.4 - Multiple probability plot of ML estimates with a fitted acceleration relationship


# 19.2.5 - Checking other model assumptions


# 19.2.6 - Estimates at use conditions


# 19.3 - FURTHER EXAMPLES

# 19.3.1 - Voltage aceleration


# 19.3.2 - Analysis of interval ALT data


# 19.3.3 - Analysis of a two-variable ALT 


# 19.4 - SOME PRACTICAL SUGGESTIONS FOR DRAWING CONCLUSIONS FROM ALT DATA

# 19.4.1 - Predicting product performance


# 19.4.2 - Drawing conclusion from accelerated test experiments


# 19.5 - OTHER KINDS OF ACCELERATED TESTS

# 19.5.1 - Continuous operation product accelerated testing


# 19.5.2 - Highly accelerated life tests


# 19.5.3 - Environmental stress testing


# 19.5.4 - Burn-in


# 19.5.5 - Environmental stress screening


# 19.6 - POTENTIAL PITFALLS OF ACCELERATED LIFE TESTING

# 19.6.1 - Pitfall 1: Multiple (unrecognized) failure modes


# 19.6.2 - Pitfall 2: Failure to properly quantify uncertainty


# 19.6.3 - Pitfall 3: Multiple time scales and degradation affected by more than one accelerating variable


# 19.6.4 - Pitfall 4: Masked failure mode


# 19.6.5 - Pitfall 5: Faulty comparison


# 19.6.6 - Pitfall 6: Accelerating variables can cause deceleration


# 19.6.7 - Pitfall 7: Untested design/production changes


# 19.6.8 - Pitfall 8: Drawing conclusion on the basis of specially build prototype test units

```{r}
HTML("<link rel='stylesheet' type='text/css' href='css/flat-slidy.css'>")
HTML("<script src='js/audiojs/audiojs/audio.min.js'></script>")
HTML("<script> audiojs.events.ready(function() {audiojs.createAll();}); </script>")
HTML("<script src='js/jkf-scroll.js'></script>")
```