---
title: "`r SMRD:::info('book')`"
subtitle: "`r SMRD:::info('chapter15')`"
author: "`r SMRD:::info('authors')`"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  slidy_presentation:
    smart: false
    fig_caption: yes
footer: "`r paste('SMRD: ', SMRD:::info('chapter15'))`"
runtime: shiny
graphics: yes
---

```{r intro, echo=FALSE,message=FALSE, warning=FALSE}
assign('dynamic', TRUE, envir = sys.frame(),immediate = TRUE)
source('scripts/R/setup.R')
```

# CHAPTER OVERVIEW

```{r}
shiny::includeCSS('scripts/css/flat-slidy.css')
shiny::includeScript(path = "scripts/js/audiojs/audiojs/audio.min.js",
                     "audiojs.events.ready(function() {audiojs.createAll();});")
shiny::includeScript("scripts/js/jkf-scroll.js")
```

## This chapter explains

- Important system reliability concepts like system structure, redundancy, nonrepairable and repairable systems, and maintainability and availability

- Basics concepts of system reliability modeling

- The distribution of system failure time as a function of individual component failure-time distributions

- Simple methods for using component test data to estimate system reliability

- Analysis of data with more that one failure mode

# 15.1 - INTRODUCTION

Systems are collections of discrete components interconnected to perform one or more tasks

System reliability is therefore a function of

1) The individual component reliabilities
    + Components can have different reliability measures for each function performed 
    + Upon failure, will a component be repaired or replaced?
    
2) How the components are connected wrt the function(s) in which they are involved
    + Will the failure cause degraded performance or total failure?
    + aka "System structure" or "Functional Architecture"
    
# 15.2 - SYSTEM STRUCTURE AND SYSTEM FAILURE PROBABILITY

# 15.2.1 - Time Dependancy of System Reliability

System failure probability $F_T(t|\underline{\theta})$ 

- Probability that the system fails before time $t$
- What defines a failure is typically defined by the user 
- Systems can have hundreds, if not thousands, of failure definitions (modes)

For a system with $i=1,...,s$ <u>__independent__</u> components 

$$F_T(t|\underline{\theta})=g\left[F_1(t|\underline{\theta}_1),...,F_2(t|\underline{\theta}_2) \right]$$

- Where the form of $g[\cdot]$ depends on how the components are arranged

# 15.2.2 - Systems with Components in Series

For systems with components arranged in series, the system fails when <u>__any__</u> component fails

```{r, echo=FALSE, fig.align='center', fig.width=10, fig.height=1.5}
par(mar = c(0,0,0,0))
plot(NA, axes = F, xlab = "", ylab = "", xlim = range(0,100), ylim = range(40,60))
points(c(10,90), c(50,50), pch = 16, cex = 2)
lines(c(10,90), c(50,50), lwd = 2)
points(rep(c(30,70),2),rep(c(50,50),2), pch = rep(c(15,0),each = 2), col = rep(c(0,1), each = 2), cex = 10)
text(rep(c(30,70),1),rep(c(50,50),1), c(1,2), cex = 2)
par(mar = c(5.1, 6.1, 3.1 ,2.1))
```

$$
\begin{aligned}
F_T(t)&=P(T\le t)=1-P(T>t)=1-P(T_1>t\cap T_2>t)\\\\
&=1-P(T_1>t)P(T_2>t)=1-(1-F_1(t|\underline{\theta}))(1-F_2(t|\underline{\theta}))\\\\
&=F_1(t|\underline{\theta})+F_2(t|\underline{\theta})-F_1(t|\underline{\theta})F_2(t|\underline{\theta})
\end{aligned}
$$

# 15.2.2 - Systems with Components in Series

For $s$ independent components

$$
\begin{aligned}
F_{_{T}}(t|\underline{\theta})&=1-\prod_{i=1}^s\left(1-F_i(t|\underline{\theta}_i)\right)\\\\
h_{_{T}}(t|\underline{\theta})&=\sum_{i=1}^s h_i(t|\underline{\theta}_i)
\end{aligned}
$$

And for $s$ iid components

$$
\begin{aligned}
F_{_{T}}(t|\underline{\theta})&=1-\left[1-F(t|\underline{\theta})\right]^s\\\\
h_{_{T}}(t|\underline{\theta})&=s \times h(t|\underline{\theta})
\end{aligned}
$$

# 15.2.2 - Systems with Components in Series

```{r, echo=FALSE, fig.align='center', fig.height=5.25}
par(family = "serif")
component.effect(parallel = FALSE)
```

# 15.2.2 - Systems with Components in Parallel

For systems with components arranged in parallel, the system fails when <u>__every__</u> component fails

```{r, echo=FALSE, fig.align='center', fig.width=10, fig.height=2.25}
par(mar = c(0,0,0,0))
plot(NA, axes = F, xlab = "", ylab = "", xlim = range(0,100), ylim = range(30,70))
points(c(10,90), c(50,50), pch = 16, cex = 2)
lines(c(10,90), c(50,50), lwd = 2)
points(rep(c(50,50),2),rep(c(50,50),2), pch = rep(c(15,0),each = 2), col = rep(c(0,1), each = 2), cex = 20)
points(rep(50,4),rep(c(35.4,64.6),2),pch = rep(c(15,0),each = 2), col = rep(c(0,1),each = 2), cex = 8)
text(rep(50,2),c(35.4,64.6), c(2,1), cex = 2)
par(mar = c(5.1, 6.1, 3.1 ,2.1))
```

$$
\begin{aligned}
F_T(t)&=P(T\le t)=P(T_1 \le t\cap T_2\le t)\\\\
&=P(T_1 \le t)P(T_2 \le t)=F_1(t|\underline{\theta})F_2(t|\underline{\theta})
\end{aligned}
$$

# 15.2.2 - Systems with Components in Series

For $s$ independent components

$$
\begin{aligned}
F_{_{T}}(t|\underline{\theta})&=\prod_{i=1}^s F_i(t|\underline{\theta}_i)\\\\
\end{aligned}
$$

And for $s$ iid components

$$
\begin{aligned}
F_{_{T}}(t|\underline{\theta})&=\left[F(t|\underline{\theta})\right]^s\\\\
\end{aligned}
$$

# 15.2.3 - Systems with Components in Parallel

```{r, echo=FALSE, fig.align='center', fig.height=5.25}
par(family  = "serif")
component.effect(parallel = TRUE)
```

# 15.2.5 - Bridge System Structure


```{r, echo=FALSE, fig.align='center', fig.width=10}
plot(NA, axes = F, xlab = "", ylab = "", xlim = range(0,100), ylim = range(0,100))
points(c(10,90), c(50,50), pch = 16, cex = 2)
lines(c(10,90), c(50,50), lwd = 2)
points(rep(c(37.6,62.4),2),rep(c(50,50),2), pch = rep(c(15,0),each = 2), col = rep(c(0,1), each = 2), cex = 30)
points(rep(50,2),rep(50,2),pch = rep(c(15,0),each = 1), col = rep(c(0,1), each = 1), cex = 8)
text(rep(50,1),rep(50,1), 3, cex = 2)
points(c(rep(37.6,4),rep(62.4,4)),rep(c(32.4,67.6),4),pch = rep(c(15,15,0,0),2), col = rep(c(0,0,1,1),2), cex = 8)
```

# 15.2.6 - $k$-out-of$s$ System Structure


# 15.3 - ESTIMATIING SYSTEM RELIABILITY FROM COMPONENT DATA

# 15.3.1 - Computing $S_T(t)$ from Component Reliability

# 15.3.2 - Sources of reliability data

# 15.3.3 - Maximum likelihood estimation of system reliability

# 15.3.4 - Normal Approximation CI for System Reliability

# 15.3.5 - Bootstrap Approximate CI for System Reliability

# 15.4 - ESTIMATING RELIABILITY WITH TWO OR MORE CAUSES OF FAILURE

# 15.4.1 - Products with two or more causes of failure

# 15.4.2 - Estimation with two or more causes of failure

# 15.4.3 - Estimation of multiple failure mode distributions when only some failure modes are identified in the data


# 15.5 - OTHER TOPICS IN SYSTEM RELIABILITY

# 15.5.1 - Other system structures

# 15.5.2 - Dependancy among components

# 15.5.3 - Systems with repair

# 15.5.4 - FMEA/FMECA

# 15.5.5 - Fault trees

# 15.5.6 - Component importance

# 15.5.7 - Markov and other state-space reliability models

```{r}

```