---
output:
  html_document:
params:
  setup: '`r SMRD.resources:::setup( )`'
  show_solution: true
---

__Show that putting together a one-sided lower and a one-sided upper $100(1-\alpha/2)\%$ prediction bounds for a future observation results in a two-sided $100(1 - \alpha)\%$ prediction interval for that observation.__

`r SMRD.resources:::solution()$begin`
A $100(1-\alpha/2)\%$ one-sided lower prediction interval for a future observation is expressed as

$$
Pr(0 \le T \le \overset{\sim}{T})=1-\alpha/2.
$$ 

This can be represented graphically (assuming a normal distribution and $\alpha = 0.05$) as shown in the plot below.

```{r, fig.align='center'}
par(lwd = 2, las = 1, yaxs = 'i')
llim <- qnorm(0.975)
curve(dnorm(x), xlim = c(-3,3), ylab = 'P(X = x)')
abline(v = llim, col = 2)
polygon(x = c(-3, seq(-3,llim, length = 100), llim),
        y = c(0, dnorm(seq(-3,llim, length = 100)), 0),
        border = F,
        col = scales::alpha('red', 0.25))
text(0,dnorm(0)/2, '0.975', cex = 1.5)
mtext(side = 3, at = llim, 
      expression(bold(tilde('T'))), 
      line = 0.005, cex = 1.1)
```

Similarly, a $100(1-\alpha/2)\%$ upper one-sided PI for a future observation is expressed as 

$$
Pr(\underset{\sim}{T} \le T < \infty)=1-\alpha/2.
$$

This can be represented graphically (assuming a normal distribution and $\alpha = 0.05$) as shown in the plot below.

```{r, fig.align='center'}
par(lwd = 2, las = 1, yaxs = 'i')
ulim <- qnorm(0.025)
curve(dnorm(x), xlim = c(-3,3), ylab = 'P(X = x)')
abline(v = ulim, col = 2)
polygon(x = c(ulim,seq(ulim, 3, length = 100), 3),
        y = c(0, dnorm(seq(ulim, 3, length = 100)), 0),
        border = F,
        col = scales::alpha('red', 0.25))
text(0,dnorm(0)/2, '0.975', cex = 1.5)
mtext(side = 3, at = ulim,'~', line = -0.250, font = 2, cex = 1.1)
mtext(side = 3, at = ulim,'T', line = 0.250, font = 2, cex = 1.1)
```

The area contained in the tails of each of the above one-sided $100(1-\alpha/2)\%$ PIs represents $\alpha/2\%$ of the total area under the curve. Combining these two one-sided PI's results in a $100(1-\alpha)%$ two-sided PI expressed as 

$$
Pr(\underset{\sim}{T} \le T < \overset{\sim}{T})=1-\alpha.
$$

The endpoints of this two-sided PI correspond to those of both of the one-sided PI's at $[\underset{\sim}{T},\overset{\sim}{T}]$. This two-sided interval is shown graphically below.

```{r, fig.align='center'}
par(lwd = 2, las = 1, yaxs = 'i')
lims <- qnorm(c(0.025, 0.975))
curve(dnorm(x), xlim = c(-3,3), ylab = 'P(X = x)')
abline(v = c(lims[1],lims[2]), col = 2)
polygon(x = c(lims[1],seq(lims[1], lims[2], length = 100), lims[2]),
        y = c(0, dnorm(seq(lims[1], lims[2], length = 100)), 0),
        border = F,
        col = scales::alpha('red', 0.25))
text(0,dnorm(0)/2, '0.950', cex = 1.5)
mtext(side = 3, at = lims[1],'~', line = -0.250, font = 2, cex = 1.1)
mtext(side = 3, at = lims[1],'T', line = 0.250, font = 2, cex = 1.1)
mtext(side = 3, at = lims[2], 
      expression(bold(tilde('T'))), 
      line = 0.005, cex = 1.1)
```

The probability of these bounds results in the total area under the curve minus the probability of the tails, $Pr(\underset{\sim}{T}\le T \le \overset{\sim}{T})=1-\alpha$, which is equivalent to a two-sided $100(1-\alpha)%$ PI.
`r SMRD.resources:::solution()$end`
