---
title: "`r SMRD:::info('book')`"
subtitle: "`r SMRD:::info('chapter11')`"
author: "`r SMRD:::info('authors')`"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  slidy_presentation:
    smart: false
    fig_caption: yes
footer: "`r paste('SMRD: ', SMRD:::info('chapter11'))`"
runtime: shiny
graphics: yes
---

```{r intro, echo=FALSE,message=FALSE, warning=FALSE}
assign('dynamic', TRUE, envir = sys.frame(),immediate = TRUE)
source('scripts/R/setup.R')
```

# CHAPTER OVERVIEW

```{r}
shiny::includeCSS('scripts/css/flat-slidy.css')
shiny::includeScript(path = "scripts/js/audiojs/audiojs/audio.min.js",
                     "audiojs.events.ready(function() {audiojs.createAll();});")
shiny::includeScript("scripts/js/jkf-scroll.js")
```

## This chapter explains


- ML estimation for th egamma and the extended generalized gamma (EGENG) distributions
- ML estimation for the Birnbaum-Saunders (BISA) and the inverse Gaussian (IGAU) distributions
- ML estimation for the limmited failure population (LFP) model
- How truncation arises in reliability data applications and ML estimation for truncated data (or data from truncated distributions)
- ML estimation for distributions with a threshold parameter like the three-parameter lognormal and the three-parameter Weibull distributions [using the generalized threshold-scale (GETS) distribution]
- Potential difficulties involved in using distributions with threshold parameters and how to avoid them

# 11.1 - Introduction

# 11.1.1 - Likelihood for other distributions and models


# 11.1.2 - CIs for other distributions and models


# 11.2 - Fitting the Gamma Distribution

# 11.3 - Fitting the Extended Generalized Gamma Distribution

# 11.4 - Fitting the BISA and IGAU Distributions

# 11.5 - Fitting a Limited Failure Population Model

# 11.5.1 - Example and data


# 11.5.2 - The limited failure population model


# 11.5.3 - The likelihood function and its maximum


# 11.5.4 - Profile likelihood functions and likelihood-based CI for $\mu,\sigma$ and $p$


# 11.6 - Truncated Data and Truncated Distributions

<font color="blue">__Censored Data__</font>

- Left censored $\in (0,t_1]$ - Event occurs prior to first observation period
- Right censored $\in [t_{m+1},\infty)$ - Event occurs after final observation period
- Interval censored $\in [t_{i-1},t_i]$ - Event occurs between observation periods 

<font color="blue">__Truncated Data__</font>

- Left truncated $\in [\tau_L,\infty)$ - Events cannot occur before $\tau_L$  
- Right truncated $\in (0,\tau_R]$ - Events cannot occur after $\tau_R$  
- Interval truncated $\in [\tau_L,\tau_R)$ - Events cannot occur before $\tau_L$ OR after $\tau_R$

A truncated distribution is "conditional" in that responses are recorded conditioned on them falling  within the truncating limits

Recall, conditioning respresents a restriction on the sample space of values that a random variable can take on 

# 11.6.1 - Examples of Left Truncation


Ultrasonic inspection is used to detect flaws in titanium alloys at key points in the manufacture of engine turbine disks

Under sustained stresses, flaws may grow to become cracks that can cause engine failure

The growth rate of a flaw/crack depends on the applied stress and the current length

- Longer cracks grow faster

Engine inspections often occur at discrete intervals 

It is very important that flaws/cracks are discovered as early as possible


During ultrasonic inspections, flaws are found by observing the amplitude of ultrasonic energy reflected by the flaw's surface

As flaw size decreases, it becomes difficult for inspectors to distinguish between a flaw and the grain structure of the material

- It is common practice for inspectors to establish material-specific thresholds $\tau_L$
- The thresholds represent the base "noise-level" of a material
- Any response observed below $\tau_L$ is discarded
- The remaining distribution of responses is said to be "left-truncated"

# 11.6.2 - Likelihood with Left Truncation


# 11.6.3 - Nonparametric estimation with left truncation


# 11.6.4 - ML estimation with left-truncated data


# 11.6.5 - Examples of right truncation


# 11.6.6 - likelihood with right (and left) truncation


# 11.6.7 - Nonparametric estimation with right (and left) truncation


# 11.7 - Fitting Distributions that have a Threshold Parameter

# 11.7.1 - Estimation with a given threshold parameter


# 11.7.2 - Probability plotting methods


# 11.7.3 - Likelihood Methods


# 11.7.4 - Summary of results of fitting models to skewed distributions

```{r}

```