---
title: "STAT 687 - Chapter 20"
subtitle: "Planning Accelerated Life Tests"
author: "Maj Jason Freels"
date: "Air Force Institute of Technology"
output: 
  ioslides_presentation:
    widescreen: true
    logo: images/afit-logo-bg.gif
    css: codes/custom.css
runtime: shiny
---

```{r echo=FALSE,message=FALSE}
foot<-"STAT 687 - Chapter 20"

connected<-FALSE

if(connected==TRUE)  {

setInternet2(TRUE)  
path   <-"https://afit.blackboard.com/bbcswebdav/users/jfreels"
mathjax<-"https://cdn.mathjax.org/mathjax/latest/MathJax.js"

download.file(paste(c(path,"DFORRT.dll"), collapse="/"),
              paste(c(getwd(),"DFORRT.dll"), collapse="/"),
              method="internal",mode="ab")

download.file(paste(c(path,"SPLIDA.dll"), collapse="/"),
              paste(c(getwd(),"SPLIDA.dll"), collapse="/"),
              method="internal",mode="ab")

dyn.load("DFORRT.dll",envir=globalenv())  ;  dyn.load("SPLIDA.dll",envir=globalenv())

a<-url(paste(c(path,"echapters.RData"    ), collapse="/")); load(a, envir=globalenv())
b<-url(paste(c(path,"RSplida Codes.RData"), collapse="/")); load(b, envir=globalenv())
c<-url(paste(c(path,"RSplidaData.RData"  ), collapse="/")); load(c, envir=globalenv())
d<-url(paste(c(path,"RSplidaUser.RData"  ), collapse="/")); load(d, envir=globalenv())
e<-url(paste(c(path,"Text data.RData"    ), collapse="/")); load(e, envir=globalenv())  }

if (connected==FALSE)  {

path   <- getwd()
mathjax<-"C:/Program Files/RStudio/resources/mathjax-23/MathJax.js"

setwd(paste(c(path,"codes"), collapse="/"))
for(i in 1:5){     load(list.files( )[grep("RData",strsplit(list.files( ),"\\W+"))][i],envir=globalenv()) } 
for(i in 1:2){ dyn.load(list.files( )[grep("dll"  ,strsplit(list.files( ),"\\W+"))][i],envir=globalenv()) }
setwd("..")            }

logo<-paste(c(path,"images","afit-logo-bg.gif"), collapse="/")
SplidaOptions(SPLIDA.DateOnPlot=FALSE)
library(survival)
library(actuar)
library(evd)
```

<script type="text/javascript" src="`r mathjax`"></script>

<script type="text/x-mathjax-config">

MathJax.Hub.Config({
      config: ["MMLorHTML.js"],
         jax: ["input/TeX","input/MathML","output/HTML-CSS","output/NativeMML"],
  extensions: ["tex2jax.js","mml2jax.js","MathMenu.js","MathZoom.js"],
         TeX: {extensions: ["AMSmath.js","AMSsymbols.js","noErrors.js","noUndefined.js"]},
  "HTML-CSS": { preferredFont: "STIX", scale: 200 }                                       });
  
</script>

<style>.foot { display: block; 
                 width: 300px; 
                height: 30px; 
               opacity: 1; 
                  left: 10%; 
                bottom: 2.15%; 
              position: absolute; 
             font-size: 60%; 
            font-color: black;}     </style>
<style>.vspace{margin-bottom: 3cm;  }     </style>

### This chapter explains
<div class="foot">`r foot`</div><IMG style="position:absolute;bottom:2.5%;LEFT:5%;WIDTH:35px;"SRC="`r logo`">

- Criteria for planning accelerated life tests (ALT)

- Simulation and analytical methods for evaluating proposed ALT plans

- The value and limitations of theoretically optimum ALT plans

- Compromis accelerated test plans that have good statistical properties and at the same time, meet practical constraints

- How to extend methods for planning single variable ALT plans to planning ALT with more than one variable 

# Section 20.1|Introduction

### $\S$ 20.1.1 - Planning information
<div class="foot">`r foot`</div><IMG style="position:absolute;bottom:2.5%;LEFT:5%;WIDTH:35px;"SRC="`r logo`">

### $\S$ 20.1.2 - Model assumptions
<div class="foot">`r foot`</div><IMG style="position:absolute;bottom:2.5%;LEFT:5%;WIDTH:35px;"SRC="`r logo`">

### $\S$ 20.1.3 - Traditional test plans
<div class="foot">`r foot`</div><IMG style="position:absolute;bottom:2.5%;LEFT:5%;WIDTH:35px;"SRC="`r logo`">

# Section 20.2|Evaluation of Test Plans

### $\S$ 20.2.1 - Evaluation using Monte Carlo simulation
<div class="foot">`r foot`</div><IMG style="position:absolute;bottom:2.5%;LEFT:5%;WIDTH:35px;"SRC="`r logo`">

### $\S$ 20.2.2 - Evaluation using large sample approximations
<div class="foot">`r foot`</div><IMG style="position:absolute;bottom:2.5%;LEFT:5%;WIDTH:35px;"SRC="`r logo`">

# Section 20.3|Planning Single-Variable ALT Experiments

### $\S$ 20.3.1 - Specifying the ALT plan
<div class="foot">`r foot`</div><IMG style="position:absolute;bottom:2.5%;LEFT:5%;WIDTH:35px;"SRC="`r logo`">

### $\S$ 20.3.2 - Planning Criteria
<div class="foot">`r foot`</div><IMG style="position:absolute;bottom:2.5%;LEFT:5%;WIDTH:35px;"SRC="`r logo`">

### $\S$ 20.3.3 - Statistically optimum test plan 
<div class="foot">`r foot`</div><IMG style="position:absolute;bottom:2.5%;LEFT:5%;WIDTH:35px;"SRC="`r logo`">

### $\S$ 20.3.4 - Compromise test plans 
<div class="foot">`r foot`</div><IMG style="position:absolute;bottom:2.5%;LEFT:5%;WIDTH:35px;"SRC="`r logo`">

# Section 20.4|Planning Two-Variable ALT Experiments

### $\S$ 20.4.1 - Two-variable ALT model
<div class="foot">`r foot`</div><IMG style="position:absolute;bottom:2.5%;LEFT:5%;WIDTH:35px;"SRC="`r logo`">

### $\S$ 20.4.2 - Examples
<div class="foot">`r foot`</div><IMG style="position:absolute;bottom:2.5%;LEFT:5%;WIDTH:35px;"SRC="`r logo`">

### $\S$ 20.4.3 - Two-variable ALT plans
<div class="foot">`r foot`</div><IMG style="position:absolute;bottom:2.5%;LEFT:5%;WIDTH:35px;"SRC="`r logo`">

### $\S$ 20.4.4 - Optimum two-variable ALT plans
<div class="foot">`r foot`</div><IMG style="position:absolute;bottom:2.5%;LEFT:5%;WIDTH:35px;"SRC="`r logo`">

### $\S$ 20.4.5 - Splitting degenerate compromise plans
<div class="foot">`r foot`</div><IMG style="position:absolute;bottom:2.5%;LEFT:5%;WIDTH:35px;"SRC="`r logo`">

### $\S$ 20.4.6 - Another example
<div class="foot">`r foot`</div><IMG style="position:absolute;bottom:2.5%;LEFT:5%;WIDTH:35px;"SRC="`r logo`">

# Section 20.5|Planning ALT Experiemnts with More Than Two Experimental Variables

```{r}
HTML("<link rel='stylesheet' type='text/css' href='css/flat-slidy.css'>")
HTML("<script src='js/audiojs/audiojs/audio.min.js'></script>")
HTML("<script> audiojs.events.ready(function() {audiojs.createAll();}); </script>")
HTML("<script src='js/jkf-scroll.js'></script>")
```